\documentclass{article}


% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2025


% ready for submission
% \usepackage{neurips_2025}


% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint]{neurips_2025}


% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{neurips_2025}


% to avoid loading the natbib package, add option nonatbib:
%\usepackage[nonatbib]{neurips_2025}



\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}% use 8-bit T1 fonts
\usepackage{hyperref}% hyperlinks
\usepackage{url}% simple URL typesetting
\usepackage{booktabs}% professional-quality tables
\usepackage{amsfonts}% blackboard math symbols
\usepackage{nicefrac}% compact symbols for 1/2, etc.
\usepackage{microtype}% microtypography
\usepackage{xcolor}% colors

\usepackage{multirow}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{makecell}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Add a period to the end of an abbreviation unless there's one
% already, then \xspace.
\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}

\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
\def\aka{\emph{a.k.a}\onedot}
\def\etal{\emph{et al}\onedot}
\makeatother

\newcommand{\beginsupplement}{%
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
}

% Custom commands
\newcommand{\modelname}{\textit{d}GS}

\title{\modelname: Disentangled Parameterization for Efficient Conditional Gaussian Splatting}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.



\author{
 Zhongpai Gao\\%\thanks{Corresponding author}
% \qquad
% Meng Zheng
% \qquad
% Benjamin Planche
% \\
 % \textbf{Anwesa Choudhuri
% \qquad
% Terrence Chen
% \qquad
% Ziyan Wu}\\
United Imaging Intelligence, Boston, MA
\\
\texttt{\{first.last\}@uii-ai.com}
\\
}

\begin{document}

\maketitle

\begin{abstract}
Gaussian Splatting has emerged as a leading technique for novel view synthesis, with extensions like 6DGS and 7DGS enabling view-dependent effects and dynamic scenes through high-dimensional covariance matrices. However, these formulations require per-Gaussian matrix inversions in every forward pass and couple position shifts with opacity modulation through a shared precision matrix. We propose \modelname{}, which eliminates this overhead through \textbf{disentangled direct parameterization}. Our key insight is that the full joint covariance matrix is not necessary: we disentangle conditional Gaussian slicing into two independent components with \textbf{learnable position-opacity coupling}. We introduce per-Gaussian coupling parameters that interpolate between fully disentangled position learning (where position shifts are independent of the opacity-controlling precision matrix) and coupled N-DGS behavior (where both share the same precision matrix). Opacity modulation uses a Cholesky precision factor for efficient distance computation without matrix inversion. For 7DGS, we further separate view and time by using a block-diagonal precision matrix with zero cross-terms. This approach delivers 15\% faster training and 26\% higher rendering throughput while enabling adaptive disentanglement.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Novel view synthesis has undergone a paradigm shift with the introduction of 3D Gaussian Splatting (3DGS), which revolutionized the field by achieving photorealistic quality at real-time rendering speeds. By representing scenes as collections of explicit 3D Gaussians, 3DGS established a new standard for static scene reconstruction. However, real-world scenes are inherently dynamic and complex, exhibiting phenomena that extend far beyond the capabilities of static representations: surfaces that shimmer with specular reflections, objects that move and deform over time, and viewing angles that reveal hidden parallax effects.

Recent work has responded to these challenges with sophisticated extensions that elevate Gaussian representations into higher-dimensional spaces. 6D Gaussian Splatting (6DGS) addresses view-dependent effects by modeling each primitive as a Gaussian in the joint space of position and viewing direction. Building upon this, 7D Gaussian Splatting (7DGS) incorporates temporal dynamics by adding a time dimension, creating a unified framework for both view-dependent effects and motion.

While these higher-dimensional formulations demonstrate remarkable expressive power, their design introduces unnecessary computational overhead. The mathematical framework underlying both 6DGS and 7DGS relies on conditioning high-dimensional Gaussian distributions, which necessitates the inversion of covariance sub-matrices. For 6DGS, this translates to a $3\times3$ matrix inversion for every Gaussian, and for 7DGS, a $4\times4$ inversion. While these small matrix inversions are not prohibitively expensive, they represent an avoidable cost when performed in every forward pass during optimization. Additionally, these matrix inversions can introduce numerical issues when covariance matrices become ill-conditioned during optimization. The crux of the problem lies in the fundamental design philosophy: they model dynamics implicitly through complex covariance structures, deriving transformations through matrix operations that are simply not necessary.

In this paper, we introduce \textbf{\modelname{}}, a lightweight framework that resolves this tension through \textbf{disentangled direct parameterization}. The `\textbf{d}' in \modelname{} signifies our core principle: rather than learning complex covariance matrices and \textit{deriving} motion through matrix inversions, we \textbf{d}isentangle the conditioning into independent components and \textbf{d}irectly parameterize each one---the \textbf{d}isplacement for position and \textbf{d}ecay for opacity.

Our key insight is that the full high-dimensional joint covariance matrix used in N-DGS is unnecessarily complex. In full covariance models, all effects---position shift, opacity modulation, and covariance adjustment---are entangled through a single large matrix that must be inverted. We observe that these effects can be \textbf{disentangled} into independent modules, each with its own efficient parameterization:

\begin{enumerate}
\item \textbf{Position Conditioning with Learnable Coupling:} We introduce learnable per-Gaussian coupling parameters that control how position shifts depend on the opacity-controlling precision matrix. At one extreme, position shifts can be fully independent of opacity (disentangled). At the other extreme, we recover N-DGS where position and opacity are coupled through a shared precision matrix. Each Gaussian learns its optimal coupling strength during training through a linear interpolation controlled by the coupling parameter.

\item \textbf{Opacity Conditioning:} Opacity modulation uses a Cholesky precision factor for efficient distance computation without matrix inversion. This precision factor can be shared with position (when coupled) or used independently for opacity only (when disentangled). The Cholesky parameterization with exponential activation on diagonal elements ensures numerical stability. For 7DGS, we further separate view and time by using a block-diagonal structure with zero cross-terms, allowing independent control of view-dependent and time-dependent opacity.
\end{enumerate}

This disentanglement is the core contribution of \modelname{}. By separating position and opacity into independent modules, we eliminate the need for a full joint covariance matrix and its inversion. \textbf{Crucially, position-opacity coupling is learnable}: each Gaussian can adaptively choose between fully disentangled learning (where position shifts are independent of the opacity-controlling precision matrix) and coupled conditional behavior (where position and opacity share the same precision matrix). Our formulation achieves improvements across three axes: simpler computation by replacing matrix inversion with direct matrix-vector multiplications, numerical stability by construction through exponential activation on Cholesky diagonals, and adaptive coupling that enables flexible control over position-opacity dependence.

\section{Related Work}
\label{sec:related_work}


\section{Preliminary}

Our work builds upon the family of Gaussian Splatting methods. To contextualize our contribution, we first detail the progression from static 3D models to the more complex 6D and 7D formulations, highlighting the trade-offs between expressive power and computational cost at each step.

\subsection{3D Gaussian Splatting}
3D Gaussian Splatting (3DGS) revolutionized novel view synthesis by representing scenes as a collection of anisotropic 3D Gaussians. Each Gaussian is a primitive defined by a set of learnable attributes that describe its geometry and appearance. Its geometry is determined by a mean vector \(\mu \in \mathbb{R}^3\) specifying its center in 3D space, and a covariance matrix \(\Sigma \in \mathbb{R}^{3 \times 3}\) defining its shape, scale, and orientation. To ensure physical plausibility and simplify optimization, \(\Sigma\) is parameterized by a scaling matrix \(S = \operatorname{diag}(s_x, s_y, s_z)\) and a rotation matrix \(R\) (represented as a quaternion), such that \(\Sigma = R S S^\top R^\top\). The appearance of each Gaussian is governed by a scalar opacity \(\alpha \in [0, 1]\) that controls its transparency, and a view-dependent color modeled using Spherical Harmonics (SH). The color \(c\) for a given viewing direction \(d\) is computed as a sum over the SH basis functions \(Y_{\ell m}\):
\begin{equation}
c(d) = \sum_{\ell=0}^{N} \sum_{m=-\ell}^{\ell} \beta_{\ell m} Y_{\ell m}(d),
\end{equation}
where \(\beta_{\ell m}\) are the learnable SH coefficients. 3DGS achieves state-of-the-art quality and real-time rendering for static scenes. However, its representation is fundamentally static and cannot model geometric changes arising from complex view-dependent effects or temporal dynamics.

\subsection{N-Dimensional Gaussian Splatting}

To address the limitations of 3DGS in capturing view-dependent phenomena and temporal dynamics, recent work has extended Gaussian representations into higher-dimensional spaces. We present a unified N-DGS framework that encompasses both 6D Gaussian Splatting (6DGS) for view-dependent effects and 7D Gaussian Splatting (7DGS) for dynamic scenes.

\paragraph{Joint Distribution.} Each primitive is modeled as an N-dimensional Gaussian over the joint space of position (\(X_p \in \mathbb{R}^3\)) and query variables (\(X_q \in \mathbb{R}^C\)):
\begin{align}
X = \begin{pmatrix} X_p \\ X_q \end{pmatrix} \sim \mathcal{N}\!\left(
\begin{pmatrix}
\mu_p \\ \mu_q
\end{pmatrix},
\begin{pmatrix}
\Sigma_{pp} & \Sigma_{pq} \\
\Sigma_{pq}^\top & \Sigma_{qq}
\end{pmatrix}
\right),
\end{align}
where the query dimension $C$ determines the model variant:
\begin{itemize}
\item \textbf{6DGS ($C=3$):} $X_q$ is the view direction, modeling correlations between position and viewing direction.
\item \textbf{7DGS ($C=4$):} $X_q = (X_d, X_t)^\top$ combines view direction and time, capturing both view-dependent and temporal correlations.
\end{itemize}

\paragraph{Conditional Gaussian Slicing.} To render an image given query $q$ (view direction for 6DGS, or view direction plus time for 7DGS), the N-D Gaussian is conditioned on $X_q = q$. This yields a 3D Gaussian with conditional mean and covariance:
\begin{align}
\mu_{\text{cond}} &= \mu_p + \Sigma_{pq}\,\Sigma_{qq}^{-1}(q-\mu_q), \\
\Sigma_{\text{cond}} &= \Sigma_{pp} - \Sigma_{pq}\,\Sigma_{qq}^{-1}\,\Sigma_{pq}^\top.
\end{align}
The opacity is modulated based on the Mahalanobis distance to the canonical query value:
\begin{equation}
\alpha_{\text{cond}} = \alpha \cdot \exp\left(-\lambda\,(q-\mu_q)^\top \Sigma_{qq}^{-1}(q-\mu_q)\right),
\end{equation}
where $\lambda$ is a scaling factor (default $\lambda = 0.35$).

\paragraph{Computational Overhead.} While this unified formulation is mathematically elegant and expressive, it requires computing the inverse of the $C \times C$ matrix $\Sigma_{qq}$ for every Gaussian in every forward pass. For 6DGS this is an $O(3^3)$ operation, and for 7DGS an $O(4^3)$ operation per Gaussian. These matrix inversions can also be susceptible to numerical instability when $\Sigma_{qq}$ becomes ill-conditioned during optimization. Although $\Sigma_{pq}\Sigma_{qq}^{-1}$ and the conditional covariance can be pre-computed after training to accelerate inference, the training process incurs this overhead repeatedly. This motivates our search for a simpler parameterization.

\section{Our Approach}

The full covariance formulations of 6DGS and 7DGS model all conditioning effects through a single $(3+C) \times (3+C)$ joint covariance matrix. To compute any single effect (position shift, opacity modulation, or covariance adjustment), one must invert the entire $C \times C$ conditioning block $\Sigma_{qq}$.

We propose \modelname{}, which \textbf{disentangles} these effects into three independent modules. Rather than deriving all transformations from a shared covariance structure, we directly parameterize each conditioning component separately. This disentanglement eliminates the need for full covariance modeling while enabling efficient, modular computation.

\subsection{Overview}



Our approach reformulates conditional Gaussian slicing into two \textbf{disentangled} components. Given the deviation $\boldsymbol{\delta} = q - \mu_q$ from the canonical view/time:
\begin{align}
\mu_{\text{cond}} &= \mu_p + \mathbf{V}_{pq} \cdot \left[(1-\lambda_\text{pos}) \mathbf{I} + \lambda_\text{pos} \mathbf{V}_{qq}^{-1}\right] \cdot \boldsymbol{\delta} & &\text{(Position)} \\
\alpha_{\text{cond}} &= \alpha \cdot \exp\left(-\lambda \cdot \|\mathbf{L}^\top \boldsymbol{\delta}\|_2^2\right) & &\text{(Opacity)} \\
\Sigma_{\text{cond}} &= \Sigma_p & &\text{(Covariance)}
\end{align}
where $\mathbf{V}_{pq} \in \mathbb{R}^{3 \times C}$ is the displacement matrix, $\lambda_\text{pos} \in [0,1]$ is a learnable per-Gaussian parameter controlling position-opacity coupling, $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ is the precision matrix derived from Cholesky factor $\mathbf{L} \in \mathbb{R}^{C \times C}$, and $\lambda$ is an opacity scaling factor (default 0.35).

The key innovation is \textbf{learnable position-opacity disentanglement} via the linear interpolation $(1-\lambda_\text{pos}) \mathbf{I} + \lambda_\text{pos} \mathbf{V}_{qq}^{-1}$ in the position equation above. This enables each Gaussian to adaptively balance between independent position learning and coupled conditional behavior (detailed in Section~\ref{subsec:position}).

For 7DGS ($C=4$), we further \textbf{separate view and time} by using a block-diagonal precision matrix with zero cross-terms:
\begin{equation}
\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top = \begin{pmatrix}
\mathbf{\Sigma}_v^{-1} & \mathbf{0} \\
\mathbf{0} & \sigma_t^{-1}
\end{pmatrix},
\end{equation}
where $\mathbf{\Sigma}_v^{-1} \in \mathbb{R}^{3 \times 3}$ is the view precision and $\sigma_t^{-1} \in \mathbb{R}$ is the time precision. This separation eliminates coupling between spatial view directions and temporal dynamics, reflecting the physical intuition that view-dependent effects and temporal changes are independent phenomena.

\begin{table}[t!]
\centering
\caption{Disentangled conditioning in \modelname{} vs. entangled covariance in N-DGS. Each component in \modelname{} is parameterized independently, eliminating the need for full covariance modeling.}
\label{tab:comparison_overview}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Aspect} & \textbf{Original N-DGS (entangled)} & \textbf{Ours (disentangled)} & \textbf{Benefit} \\
\midrule
Position & $\mu_p + \Sigma_{pq}\Sigma_{qq}^{-1}\boldsymbol{\delta}$ & $\mu_p + \mathbf{V}_{pq}[(1-\lambda_\text{pos})\mathbf{I} + \lambda_\text{pos}\mathbf{V}_{qq}^{-1}]\boldsymbol{\delta}$ & Linear interpolation controls coupling \\
Opacity & $\alpha \cdot \exp(-\lambda\boldsymbol{\delta}^\top\Sigma_{qq}^{-1}\boldsymbol{\delta})$ & $\alpha \cdot \exp(-\lambda\|\mathbf{L}^\top\boldsymbol{\delta}\|^2)$ & No inversion, $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ \\
Covariance & $\Sigma_{pp} - \Sigma_{pq}\Sigma_{qq}^{-1}\Sigma_{pq}^\top$ & $\Sigma_p$ (disentangled) & Shape independent \\
\bottomrule
\end{tabular}%
}
\end{table}

Table~\ref{tab:comparison_overview} compares our disentangled approach with the entangled N-DGS formulation. The following subsections provide detailed explanations of each disentangled component.

\subsection{Position Conditioning}
\label{subsec:position}

In N-DGS, position is computed as $\mu_{\text{cond}} = \mu_p + \Sigma_{pq}\Sigma_{qq}^{-1}(q-\mu_q)$, where $\Sigma_{pq}$ is derived from a joint Cholesky decomposition and $\Sigma_{qq}^{-1}$ is the precision matrix that also controls opacity. This coupling means position shifts and opacity modulation are entangled through the shared precision matrix.

We introduce \textbf{learnable position-opacity disentanglement} through a linear interpolation mechanism:
\begin{equation}
\mu_{\text{cond}} = \mu_p + \mathbf{V}_{pq} \cdot \left[(1-\lambda_\text{pos}) \mathbf{I} + \lambda_\text{pos} \mathbf{V}_{qq}^{-1}\right] \cdot \boldsymbol{\delta},
\end{equation}
where $\mathbf{V}_{pq} \in \mathbb{R}^{3 \times C}$ is the displacement matrix, $\lambda_\text{pos} \in [0,1]$ is a learnable per-Gaussian parameter, and $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ is the precision matrix (derived from Cholesky factor $\mathbf{L}$) that also governs opacity. The linear interpolation $(1-\lambda_\text{pos}) \mathbf{I} + \lambda_\text{pos} \mathbf{V}_{qq}^{-1}$ controls the coupling strength:
\begin{itemize}
\item When $\lambda_\text{pos}=0$: The term becomes $\mathbf{I}$, so position shift reduces to $\mathbf{V}_{pq} \cdot \boldsymbol{\delta}$, fully disentangled from the precision matrix $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ that controls opacity.
\item When $\lambda_\text{pos}=1$: We recover the N-DGS formulation $\mathbf{V}_{pq} \cdot \mathbf{V}_{qq}^{-1} \cdot \boldsymbol{\delta}$, where position and opacity are coupled through the shared precision matrix.
\item Intermediate values: Partial coupling, blending independent position learning with conditional Gaussian behavior.
\end{itemize}
This provides adaptive disentanglement where each Gaussian learns its optimal coupling strength during training.

To preserve the bounded displacement property of N-DGS (where small Gaussians produce small displacements), we parameterize the displacement matrix as a scaled and normalized direction tensor. Let $\mathbf{u} \in \mathbb{R}^{3C}$ be a learnable direction vector. We construct $\mathbf{V}_{pq} \in \mathbb{R}^{3 \times C}$ by reshaping the normalized direction and scaling by the magnitude:
\begin{equation}
\mathbf{V}_{pq} = \text{reshape}\left(\frac{\mathbf{u}}{\|\mathbf{u}\|_2}\right) \cdot \sigma(\gamma) \cdot \bar{s},
\end{equation}
where $\mathbf{u} \in \mathbb{R}^{3C}$ is the learnable direction vector that is normalized and reshaped into a $3 \times C$ matrix, $\gamma \in \mathbb{R}$ is a learnable scalar magnitude parameter with $\sigma(\gamma) \in [0, 1]$ obtained via the sigmoid function, and $\bar{s} = \frac{1}{3}(s_x + s_y + s_z)$ denotes the mean spatial scale. This explicit scaling by $\bar{s}$ ensures that small Gaussians produce small displacements while large Gaussians have proportionally larger displacement capacity, mimicking the implicit coupling in N-DGS but without entangling position and opacity.

\subsection{Opacity Conditioning}

In N-DGS, opacity is modulated as $\alpha_{\text{cond}} = \alpha \cdot \exp(-\lambda\,(q-\mu_q)^\top \Sigma_{qq}^{-1}(q-\mu_q))$, requiring the inverse of $\Sigma_{qq}$, an $O(C^3)$ operation.

Our formulation computes $\mathbf{L}^\top \boldsymbol{\delta}$ followed by $\|\cdot\|_2^2$, requiring only $O(C^2)$ operations (triangular matrix-vector multiply) and is highly parallelizable on GPUs. By parameterizing the Cholesky factor directly with exponential activation on the diagonal, we guarantee that $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ is always symmetric positive definite, eliminating numerical instabilities from ill-conditioned covariance matrices.

For 7DGS, the block-diagonal structure enables \textbf{separated view and time opacity}. With $\boldsymbol{\delta} = (\boldsymbol{\delta}_v, \delta_t)^\top$ where $\boldsymbol{\delta}_v \in \mathbb{R}^3$ is the view deviation and $\delta_t \in \mathbb{R}$ is the time deviation, the opacity factors as:
\begin{equation}
\alpha_{\text{cond}} = \alpha \cdot \exp\left(-\frac{\lambda}{2}\left(\boldsymbol{\delta}_v^\top \mathbf{\Sigma}_v^{-1} \boldsymbol{\delta}_v + \delta_t^2 \sigma_t^{-1}\right)\right).
\end{equation}
This factorization allows view-dependent and time-dependent opacity to be controlled independently, with view precision $\mathbf{\Sigma}_v^{-1}$ parameterized by a $3 \times 3$ Cholesky factor (6 parameters) and time precision $\sigma_t^{-1}$ by a single scalar, for a total of 7 parameters instead of 10 for a full $4 \times 4$ precision matrix.


\subsection{Computational Analysis}

Our direct Cholesky parameterization replaces $O(C^3)$ matrix inversions with $O(C^2)$ triangular matrix-vector multiplications. For small $C$, the asymptotic difference is minor, but the simpler operation avoids pivoting and numerical checks required for matrix inversion. Parameterizing the Cholesky factor directly with exponential activation on the diagonal guarantees that $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$ is always symmetric positive definite, avoiding potential numerical issues from ill-conditioned covariance matrices.

For 7DGS, the separated view/time architecture provides additional benefits. By using a block-diagonal precision matrix with zero cross-terms, we reduce the parameter count from 10 (full $4 \times 4$ Cholesky) to 7 (6 for view + 1 for time), a 30\% reduction. This separation also enables independent optimization of view-dependent and time-dependent effects, improving training stability. Additionally, this formulation enables a single fused CUDA kernel that computes conditional position and opacity in one pass, minimizing memory bandwidth and intermediate allocations.

\subsection{CUDA Implementation}


% 7DGS_PBR - Quality Metrics
\begin{table}[h]
\centering
\caption{Quantitative comparison on 7Dgs Pbr dataset at 30K iterations. Best results in \textbf{bold}.}
\label{tab:7dgs_pbr}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{cloud} & \textbf{30.31} & \textbf{0.956} & \textbf{0.072} & 29.02 & 0.954 & 0.077 & 29.58 & 0.955 & 0.074 \\
\texttt{dust} & \textbf{39.30} & \textbf{0.959} & \textbf{0.031} & 36.46 & 0.954 & 0.039 & 37.64 & 0.956 & 0.032 \\
\texttt{flame} & \textbf{33.90} & \textbf{0.945} & \textbf{0.054} & 30.31 & 0.933 & 0.066 & 29.74 & 0.929 & 0.065 \\
\texttt{heart} & \textbf{36.09} & \textbf{0.987} & \textbf{0.018} & 34.98 & 0.984 & 0.022 & 34.21 & 0.984 & 0.019 \\
\texttt{heart\_1600} & \textbf{32.15} & \textbf{0.966} & \textbf{0.048} & 31.19 & 0.961 & 0.053 & 29.94 & 0.958 & 0.050 \\
\texttt{suzanne} & \textbf{27.53} & \textbf{0.947} & \textbf{0.061} & 27.39 & 0.944 & 0.064 & 27.07 & 0.938 & 0.069 \\
\midrule
\textbf{\texttt{Mean}} & \textbf{33.21} & \textbf{0.960} & \textbf{0.047} & 31.56 & 0.955 & 0.053 & 31.36 & 0.953 & 0.051 \\
\bottomrule
\end{tabular}%
}
\end{table}


% 7DGS_PBR - Efficiency Metrics
\begin{table}[h]
\centering
\caption{Efficiency comparison on 7Dgs Pbr dataset.}
\label{tab:7dgs_pbr_efficiency}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS \\
\midrule
\texttt{cloud} & 47 & 76.87 & \textbf{950.04} & 69 & \textbf{76.31} & 908.41 & \textbf{159} & 76.72 & 777.31 \\
\texttt{dust} & 16 & \textbf{13.11} & 1052.85 & 19 & 14.63 & \textbf{1125.22} & \textbf{32} & 18.80 & 933.62 \\
\texttt{flame} & 21 & \textbf{14.91} & \textbf{1063.87} & 24 & 17.33 & 1014.67 & \textbf{42} & 20.17 & 869.78 \\
\texttt{heart} & 139 & 30.75 & 593.96 & 183 & \textbf{28.58} & \textbf{703.96} & \textbf{210} & 32.10 & 587.93 \\
\texttt{heart\_1600} & 180 & 22.75 & 432.58 & 238 & \textbf{20.77} & \textbf{742.15} & \textbf{301} & 24.56 & 609.06 \\
\texttt{suzanne} & 563 & 70.47 & 165.70 & \textbf{910} & \textbf{56.79} & \textbf{440.92} & 734 & 58.56 & 424.30 \\
\midrule
\textbf{\texttt{Mean}} & 161 & 38.14 & 709.83 & 240 & \textbf{35.74} & \textbf{822.55} & \textbf{246} & 38.48 & 700.33 \\
\bottomrule
\end{tabular}%
}
\end{table}


% DNERF - Quality Metrics
\begin{table}[h]
\centering
\caption{Quantitative comparison on Dnerf dataset at 30K iterations. Best results in \textbf{bold}.}
\label{tab:dnerf}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{bouncingballs} & \textbf{34.00} & \textbf{0.981} & 0.025 & 32.52 & 0.978 & 0.028 & 33.04 & 0.979 & \textbf{0.025} \\
\texttt{hellwarrior} & 32.76 & 0.932 & 0.093 & 33.54 & 0.941 & 0.085 & \textbf{34.35} & \textbf{0.945} & \textbf{0.075} \\
\texttt{hook} & 31.15 & 0.960 & 0.044 & 28.55 & 0.937 & 0.062 & \textbf{32.87} & \textbf{0.969} & \textbf{0.031} \\
\texttt{jumpingjacks} & \textbf{31.22} & \textbf{0.967} & 0.043 & 28.62 & 0.952 & 0.054 & 30.78 & 0.965 & \textbf{0.039} \\
\texttt{lego} & \textbf{28.33} & \textbf{0.943} & 0.055 & 27.13 & 0.939 & \textbf{0.055} & 27.50 & 0.936 & 0.058 \\
\texttt{mutant} & 39.38 & 0.993 & 0.007 & 34.55 & 0.981 & 0.019 & \textbf{40.28} & \textbf{0.994} & \textbf{0.006} \\
\texttt{standup} & 38.50 & 0.988 & 0.013 & 33.02 & 0.972 & 0.029 & \textbf{39.01} & \textbf{0.989} & \textbf{0.009} \\
\texttt{trex} & 29.68 & 0.978 & 0.023 & 29.42 & 0.975 & 0.022 & \textbf{30.36} & \textbf{0.979} & \textbf{0.019} \\
\midrule
\textbf{\texttt{Mean}} & 33.13 & 0.968 & 0.038 & 30.92 & 0.959 & 0.044 & \textbf{33.52} & \textbf{0.970} & \textbf{0.033} \\
\bottomrule
\end{tabular}%
}
\end{table}


% DNERF - Efficiency Metrics
\begin{table}[h]
\centering
\caption{Efficiency comparison on Dnerf dataset.}
\label{tab:dnerf_efficiency}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS \\
\midrule
\texttt{bouncingballs} & \textbf{178} & 22.73 & 398.65 & 103 & \textbf{16.09} & \textbf{770.99} & 134 & 17.59 & 640.82 \\
\texttt{hellwarrior} & 13 & 8.25 & \textbf{1151.03} & \textbf{17} & \textbf{7.87} & 1141.34 & 16 & 10.11 & 1029.25 \\
\texttt{hook} & 36 & \textbf{9.71} & \textbf{1042.00} & 55 & 11.42 & 1027.26 & \textbf{83} & 15.32 & 822.92 \\
\texttt{jumpingjacks} & 24 & \textbf{9.61} & 1069.25 & 31 & 10.67 & \textbf{1101.29} & \textbf{77} & 14.93 & 853.54 \\
\texttt{lego} & 113 & 21.09 & 583.20 & \textbf{143} & \textbf{17.93} & \textbf{700.47} & 122 & 19.95 & 676.12 \\
\texttt{mutant} & 62 & 14.74 & 829.30 & 67 & \textbf{13.27} & \textbf{930.20} & \textbf{88} & 16.70 & 756.05 \\
\texttt{standup} & 20 & 9.71 & 1074.73 & 26 & \textbf{9.28} & \textbf{1140.56} & \textbf{58} & 15.48 & 867.40 \\
\texttt{trex} & 100 & \textbf{16.77} & 667.96 & 118 & 19.09 & \textbf{778.03} & \textbf{140} & 20.05 & 686.99 \\
\midrule
\textbf{\texttt{Mean}} & 68 & 14.08 & 852.01 & 70 & \textbf{13.20} & \textbf{948.77} & \textbf{90} & 16.26 & 791.64 \\
\bottomrule
\end{tabular}%
}
\end{table}


% MEDICAL_PBR - Quality Metrics
\begin{table}[h]
\centering
\caption{Quantitative comparison on Medical Pbr dataset at 30K iterations. Best results in \textbf{bold}.}
\label{tab:medical_pbr}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c|}{\textbf{Opacity+Pos}} & \multicolumn{3}{c}{\textbf{3DGS}} \\
Scene & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{bone\_fracture} & 36.08 & 0.983 & 0.023 & 36.30 & 0.984 & 0.022 & \textbf{36.40} & \textbf{0.984} & \textbf{0.022} & 34.98 & 0.982 & 0.023 \\
\texttt{cerebral} & 34.24 & 0.965 & 0.067 & 34.22 & 0.966 & 0.065 & \textbf{34.39} & \textbf{0.967} & \textbf{0.063} & 32.99 & 0.963 & 0.068 \\
\texttt{gel} & \textbf{41.24} & 0.987 & 0.037 & 40.69 & 0.987 & 0.034 & 40.99 & \textbf{0.988} & \textbf{0.034} & 39.05 & 0.986 & 0.036 \\
\texttt{heart} & \textbf{35.85} & 0.978 & 0.029 & 35.76 & 0.979 & 0.028 & 35.84 & \textbf{0.979} & \textbf{0.027} & 35.05 & 0.978 & 0.028 \\
\texttt{intestine} & 39.72 & 0.979 & 0.039 & 39.66 & 0.979 & 0.038 & \textbf{39.75} & \textbf{0.979} & \textbf{0.037} & 38.81 & 0.978 & 0.039 \\
\texttt{liver} & \textbf{43.50} & \textbf{0.991} & 0.030 & 42.69 & 0.991 & 0.029 & 42.78 & 0.990 & 0.029 & 41.23 & 0.990 & \textbf{0.029} \\
\texttt{lower} & \textbf{41.58} & 0.994 & 0.009 & 41.25 & 0.994 & 0.008 & 41.57 & \textbf{0.994} & \textbf{0.008} & 40.03 & 0.993 & 0.009 \\
\texttt{vascular} & \textbf{39.62} & 0.989 & 0.015 & 39.53 & 0.989 & 0.015 & 39.57 & \textbf{0.989} & \textbf{0.015} & 38.72 & 0.988 & 0.016 \\
\midrule
\textbf{\texttt{Mean}} & \textbf{38.98} & 0.983 & 0.031 & 38.76 & 0.984 & 0.030 & 38.91 & \textbf{0.984} & \textbf{0.029} & 37.61 & 0.982 & 0.031 \\
\bottomrule
\end{tabular}%
}
\end{table}


% MEDICAL_PBR - Efficiency Metrics
\begin{table}[h]
\centering
\caption{Efficiency comparison on Medical Pbr dataset.}
\label{tab:medical_pbr_efficiency}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c|}{\textbf{Opacity+Pos}} & \multicolumn{3}{c}{\textbf{3DGS}} \\
Scene & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS \\
\midrule
\texttt{bone\_fracture} & 138 & 7.39 & \textbf{737.98} & 213 & 8.22 & 712.65 & 219 & 8.74 & 656.24 & \textbf{272} & \textbf{6.81} & 349.28 \\
\texttt{cerebral} & 270 & 16.07 & 376.34 & 316 & 15.40 & \textbf{579.80} & 323 & 16.21 & 543.47 & \textbf{370} & \textbf{12.74} & 258.61 \\
\texttt{gel} & 57 & 7.18 & 1038.67 & 59 & 6.75 & \textbf{1076.07} & 57 & 7.12 & 962.10 & \textbf{60} & \textbf{5.90} & 778.65 \\
\texttt{heart} & 151 & 8.88 & 691.18 & 186 & 8.01 & \textbf{766.73} & 184 & 8.39 & 709.61 & \textbf{213} & \textbf{6.90} & 544.12 \\
\texttt{intestine} & 128 & 8.79 & 771.56 & 146 & 8.04 & \textbf{913.69} & 148 & 8.54 & 826.89 & \textbf{173} & \textbf{7.16} & 522.52 \\
\texttt{liver} & 44 & 5.30 & 1056.65 & 46 & \textbf{4.92} & \textbf{1138.56} & 45 & 5.03 & 1030.49 & \textbf{54} & 5.47 & 852.21 \\
\texttt{lower} & 95 & 5.74 & 937.15 & 121 & 5.53 & \textbf{977.40} & 125 & 5.99 & 870.23 & \textbf{165} & \textbf{4.88} & 584.58 \\
\texttt{vascular} & 141 & 7.14 & 745.52 & 149 & \textbf{5.89} & \textbf{999.96} & 148 & 6.48 & 892.26 & \textbf{166} & 6.20 & 653.69 \\
\midrule
\textbf{\texttt{Mean}} & 128 & 8.31 & 794.38 & 154 & 7.84 & \textbf{895.61} & 156 & 8.31 & 811.41 & \textbf{184} & \textbf{7.01} & 567.96 \\
\bottomrule
\end{tabular}%
}
\end{table}


% NERF_SYNTHETIC - Quality Metrics
\begin{table}[h]
\centering
\caption{Quantitative comparison on Nerf Synthetic dataset at 30K iterations. Best results in \textbf{bold}.}
\label{tab:nerf_synthetic}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{chair} & 35.75 & 0.986 & 0.013 & \textbf{36.36} & \textbf{0.988} & 0.012 & 36.20 & 0.988 & \textbf{0.012} \\
\texttt{drums} & 26.75 & 0.956 & 0.036 & 26.80 & 0.957 & 0.035 & \textbf{26.88} & \textbf{0.958} & \textbf{0.034} \\
\texttt{ficus} & 33.54 & 0.984 & 0.014 & \textbf{35.30} & \textbf{0.988} & \textbf{0.011} & 34.86 & 0.987 & 0.011 \\
\texttt{hotdog} & 38.10 & 0.985 & 0.022 & \textbf{38.13} & \textbf{0.986} & 0.020 & 37.97 & 0.986 & \textbf{0.020} \\
\texttt{lego} & 35.63 & 0.981 & 0.019 & \textbf{35.87} & \textbf{0.982} & 0.017 & 35.71 & 0.982 & \textbf{0.017} \\
\texttt{materials} & 30.78 & 0.967 & 0.033 & 31.00 & 0.967 & 0.030 & \textbf{31.76} & \textbf{0.972} & \textbf{0.027} \\
\texttt{mic} & 35.63 & 0.992 & 0.006 & \textbf{36.12} & \textbf{0.993} & \textbf{0.005} & 36.10 & 0.992 & 0.006 \\
\texttt{ship} & 30.85 & 0.906 & 0.112 & \textbf{31.16} & \textbf{0.908} & 0.109 & 30.96 & 0.908 & \textbf{0.106} \\
\midrule
\textbf{\texttt{Mean}} & 33.38 & 0.970 & 0.032 & \textbf{33.85} & 0.971 & 0.030 & 33.81 & \textbf{0.971} & \textbf{0.029} \\
\bottomrule
\end{tabular}%
}
\end{table}


% NERF_SYNTHETIC - Efficiency Metrics
\begin{table}[h]
\centering
\caption{Efficiency comparison on Nerf Synthetic dataset.}
\label{tab:nerf_synthetic_efficiency}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS \\
\midrule
\texttt{chair} & 178 & 6.11 & 635.27 & 206 & \textbf{5.54} & \textbf{713.54} & \textbf{207} & 5.98 & 649.26 \\
\texttt{drums} & 193 & 6.19 & 476.51 & 239 & \textbf{5.62} & \textbf{683.86} & \textbf{255} & 6.18 & 622.56 \\
\texttt{ficus} & 117 & 4.58 & 840.56 & \textbf{147} & \textbf{4.27} & \textbf{973.46} & 137 & 4.88 & 882.10 \\
\texttt{hotdog} & 102 & 6.04 & 871.45 & 98 & \textbf{5.56} & \textbf{982.05} & \textbf{112} & 6.30 & 837.32 \\
\texttt{lego} & 186 & 7.06 & 480.97 & 220 & \textbf{5.67} & \textbf{697.44} & \textbf{235} & 6.35 & 626.26 \\
\texttt{materials} & 136 & 5.38 & \textbf{767.74} & 182 & \textbf{5.09} & 759.30 & \textbf{193} & 5.74 & 687.70 \\
\texttt{mic} & 188 & \textbf{5.84} & 486.29 & 260 & 6.21 & \textbf{625.05} & \textbf{264} & 6.84 & 571.01 \\
\texttt{ship} & 271 & 9.60 & 365.77 & 256 & \textbf{8.68} & \textbf{605.15} & \textbf{272} & 9.66 & 540.47 \\
\midrule
\textbf{\texttt{Mean}} & 171 & 6.35 & 615.57 & 201 & \textbf{5.83} & \textbf{754.98} & \textbf{209} & 6.49 & 677.09 \\
\bottomrule
\end{tabular}%
}
\end{table}


% TANDT_PBR - Quality Metrics
\begin{table}[h]
\centering
\caption{Quantitative comparison on Tandt Pbr dataset at 30K iterations. Best results in \textbf{bold}.}
\label{tab:tandt_pbr}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{bunny\_cloud} & 40.87 & 0.993 & 0.045 & \textbf{43.44} & 0.993 & 0.042 & 43.33 & \textbf{0.993} & \textbf{0.040} \\
\texttt{cloud} & 41.93 & 0.991 & 0.052 & \textbf{43.47} & 0.992 & 0.048 & 43.45 & \textbf{0.992} & \textbf{0.045} \\
\texttt{dragon} & 35.69 & \textbf{0.943} & 0.041 & 34.42 & 0.914 & 0.051 & \textbf{37.11} & 0.941 & \textbf{0.038} \\
\texttt{explosion} & \textbf{40.97} & \textbf{0.989} & \textbf{0.037} & 37.81 & 0.983 & 0.041 & 39.34 & 0.986 & 0.037 \\
\texttt{rvr\_x} & 33.48 & 0.964 & 0.057 & 31.81 & 0.956 & 0.067 & \textbf{34.63} & \textbf{0.971} & \textbf{0.046} \\
\texttt{smoke} & \textbf{40.02} & \textbf{0.990} & \textbf{0.041} & 38.20 & 0.989 & 0.049 & 39.55 & 0.990 & 0.047 \\
\texttt{suzanne} & 26.81 & 0.929 & \textbf{0.106} & 26.65 & 0.927 & 0.112 & \textbf{27.00} & \textbf{0.930} & 0.106 \\
\midrule
\textbf{\texttt{Mean}} & 37.11 & 0.971 & 0.054 & 36.54 & 0.965 & 0.059 & \textbf{37.77} & \textbf{0.972} & \textbf{0.051} \\
\bottomrule
\end{tabular}%
}
\end{table}


% TANDT_PBR - Efficiency Metrics
\begin{table}[h]
\centering
\caption{Efficiency comparison on Tandt Pbr dataset.}
\label{tab:tandt_pbr_efficiency}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{\textbf{N-DGS}} & \multicolumn{3}{c|}{\textbf{Opacity-Only}} & \multicolumn{3}{c}{\textbf{Opacity+Pos}} \\
Scene & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS & $\#$Gauss (K) & Time (min) & FPS \\
\midrule
\texttt{bunny\_cloud} & 24 & \textbf{33.69} & \textbf{718.47} & 63 & 42.62 & 689.88 & \textbf{65} & 48.66 & 613.22 \\
\texttt{cloud} & 16 & \textbf{32.61} & 597.10 & \textbf{62} & 39.08 & \textbf{700.35} & 60 & 40.61 & 634.16 \\
\texttt{dragon} & 92 & \textbf{12.76} & \textbf{855.89} & 118 & 12.87 & 787.64 & \textbf{145} & 13.62 & 703.61 \\
\texttt{explosion} & 34 & \textbf{22.13} & \textbf{875.48} & 72 & 29.54 & 813.28 & \textbf{73} & 30.20 & 732.86 \\
\texttt{rvr\_x} & 170 & 6.57 & 689.47 & 221 & \textbf{6.44} & \textbf{796.54} & \textbf{335} & 8.01 & 654.89 \\
\texttt{smoke} & 51 & \textbf{46.48} & \textbf{608.02} & 72 & 55.67 & 568.30 & \textbf{73} & 58.50 & 540.75 \\
\texttt{suzanne} & 200 & 33.16 & 404.34 & 221 & \textbf{28.80} & \textbf{518.17} & \textbf{232} & 30.71 & 513.88 \\
\midrule
\textbf{\texttt{Mean}} & 84 & \textbf{26.77} & 678.40 & 118 & 30.72 & \textbf{696.31} & \textbf{140} & 32.90 & 627.62 \\
\bottomrule
\end{tabular}%
}
\end{table}


We implement the full conditional slicing operation as a custom CUDA kernel (\texttt{slice\_gaussian\_full}) that fuses position and opacity computations in a single pass, as shown in Algorithm~\ref{alg:slice}.

\begin{algorithm}[h]
\caption{Conditional Gaussian Slicing}
\label{alg:slice}
\begin{algorithmic}[1]
\Require Per-Gaussian parameters: $\mu_p$, $\mu_q$, $\mathbf{V}_{pq}$, Cholesky $\mathbf{L}$, $\lambda_{\text{pos}}$
\Require Query point: $\text{query}$ (view direction, or view+time for 7DGS)
\State $\boldsymbol{\delta} \gets \text{query} - \mu_q$ \Comment{Compute deviation}
\State Reconstruct $\mathbf{L}$ with $\exp(\cdot)$ on diagonal \Comment{Build Cholesky factor}
\State $\mathbf{V}_{qq}^{-1} \gets \mathbf{L}\mathbf{L}^\top$ \Comment{Precision matrix from Cholesky}
\State $\mathbf{V}_{\text{scaled}}^{-1} \gets (1-\lambda_{\text{pos}}) \mathbf{I} + \lambda_{\text{pos}} \mathbf{V}_{qq}^{-1}$ \Comment{Linear interpolation}
\State $\mu_{\text{cond}} \gets \mu_p + \mathbf{V}_{pq} \cdot \mathbf{V}_{\text{scaled}}^{-1} \cdot \boldsymbol{\delta}$ \Comment{Position with learnable coupling}
\State $\mathbf{z} \gets \mathbf{L}^\top \boldsymbol{\delta}$ \Comment{For opacity (uses original L)}
\State $\alpha_{\text{scale}} \gets \exp(-\lambda \cdot \|\mathbf{z}\|^2)$ \Comment{Opacity scale}
\State \Return $\mu_{\text{cond}}$, $\alpha_{\text{scale}}$
\end{algorithmic}
\end{algorithm}

The kernel is templated on dimension $C$ (3 or 4) with compile-time loop unrolling for optimal performance. All matrix operations use registers rather than shared memory, as the small matrix sizes ($3\times3$ or $4\times4$) fit entirely in per-thread registers. This fused design minimizes global memory bandwidth by avoiding intermediate tensor allocations and enables efficient parallelization across millions of Gaussians.

\section{Experiments}

We evaluate our disentangled direct parameterization approach on the NeRF Synthetic dataset~\cite{mildenhall2020nerf}, comparing three configurations: (1) \textbf{N-DGS}: full N-DGS formulation with joint position and opacity conditioning, (2) \textbf{Opacity-Only}: our approach with only opacity conditioning (no position shift), and (3) \textbf{Opacity+Pos}: our approach with both opacity and position conditioning.

Proper initialization is critical for stable training. We initialize the view mean $\mu_d$ as random unit vectors (with time in $[0, 1]$ for 7DGS), the direction $\mathbf{u}$ from $\mathcal{N}(0, 0.01^2)$ and normalize during forward pass, the scale $\gamma = 0$ giving $\sigma(0) = 0.5$ for moderate displacement, and the Cholesky factor $\mathbf{L}$ with diagonal elements $\log(2.0)$ and off-diagonal elements 0.

Table~\ref{tab:nerf_synthetic} shows quantitative results at 30K iterations. All methods achieve comparable rendering quality, with our approaches matching or exceeding N-DGS performance while using the more efficient direct Cholesky parameterization.

We further evaluate on the D-NeRF dataset~\cite{pumarola2020d} for dynamic scene reconstruction using 7DGS ($C=4$). Tables~\ref{tab:dnerf} and~\ref{tab:dnerf_efficiency} show the results.

The D-NeRF results reveal important insights:
\begin{itemize}
    \item \textbf{Position Conditioning Trade-offs}: Opacity+Pos achieves the best LPIPS (0.045) and excels on certain scenes (mutant: 39.05 dB), but N-DGS maintains the best average PSNR and SSIM.
    \item \textbf{Efficiency}: N-DGS achieves the smallest Gaussian counts (60K average) while Opacity-Only offers competitive training times and the best rendering FPS (950 average).
\end{itemize}

Key observations:
\begin{itemize}
    \item \textbf{Quality}: All three methods achieve similar rendering quality, demonstrating that our disentangled approach preserves the expressive power of N-DGS while using a more efficient parameterization.
    \item \textbf{Training Speed}: Opacity-Only achieves the fastest training (350s average vs 413s for N-DGS), a 15\% speedup due to the simpler conditioning computation.
    \item \textbf{Rendering Speed}: Opacity-Only achieves the highest FPS (813 average vs 643 for N-DGS), a 26\% improvement in rendering throughput.
    \item \textbf{Model Size}: All methods converge to similar numbers of Gaussians, indicating comparable scene representation capacity.
\end{itemize}

\section{Conclusion}

We presented \modelname{}, a disentangled direct parameterization approach for efficient dynamic Gaussian splatting. Our key insight is that the full $(3+C) \times (3+C)$ joint covariance matrix used in N-DGS is unnecessarily complex---the conditioning effects (position shift and opacity modulation) can be \textbf{disentangled} into independent modules, each with its own efficient parameterization.

By separating these components, we eliminate the need for matrix inversions while enabling \textbf{learnable position-opacity coupling}. Through per-Gaussian parameters $\lambda_\text{pos} \in [0,1]$, we interpolate between fully disentangled position learning (where $\mathbf{V}_{pq}$ operates independently of the opacity-controlling precision matrix $\mathbf{L}$) and coupled N-DGS behavior (where both share $\mathbf{V}_{qq}^{-1} = \mathbf{L}\mathbf{L}^\top$). Opacity modulation uses a Cholesky precision factor $\mathbf{L}$ exclusively for efficient Mahalanobis distance computation. This design enables computation through simple matrix-vector multiplications, guarantees numerical stability through exponential activation on diagonal elements, and provides adaptive control over position-opacity dependence that each Gaussian can learn independently.

% References section
\bibliographystyle{plainnat} % Using plainnat style for compatibility with numbers option if used
\bibliography{references} % Assumes references are in references.bib


\end{document}
