%%%%%%%% ICML 2026 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2026} with \usepackage[nohyperref]{icml2026} above.
\usepackage{hyperref}


% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2026}

% For preprint, use
% \usepackage[preprint]{icml2026}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2026}


% Additional packages (avoiding duplicates with template packages above)
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}% use 8-bit T1 fonts
\usepackage{url}% simple URL typesetting
\usepackage{amsfonts}% blackboard math symbols
\usepackage{nicefrac}% compact symbols for 1/2, etc.
\usepackage{xcolor}% colors

\usepackage{multirow}
\usepackage{rotating}
\usepackage{siunitx}
\usepackage{makecell}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{gensymb}
% \usepackage{algorithm}
% \usepackage[noend]{algpseudocode}
\newcommand{\return}{\textbf{return }}
\newcommand{\Comment}[1]{\hfill $\triangleright$ #1}
\usepackage{enumitem}

% Add a period to the end of an abbreviation unless there's one
% already, then \xspace.
\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}

\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
\def\aka{\emph{a.k.a}\onedot}
\def\etal{\emph{et al}\onedot}
\makeatother

\newcommand{\beginsupplement}{%
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
}

% Custom commands
\newcommand{\modelname}{\textit{d}GS}

% Note: \onedot, abbreviation commands (\eg, \ie, \etc, \etal, \wrt, \dof, \aka)
% are already defined earlier in this file
% Note: amsmath, amssymb, mathtools, amsthm already loaded above (lines 44-47)

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{\modelname: Direct Parameterization for Efficient Conditional Gaussian Splatting}

\begin{document}

\twocolumn[
  \icmltitle{\modelname: Direct Parameterization for Efficient Conditional Gaussian Splatting}

  % It is OKAY to include author information, even for blind submissions: the
  % style file will automatically remove it for you unless you've provided
  % the [accepted] option to the icml2026 package.

  % List of affiliations: The first argument should be a (short) identifier you
  % will use later to specify author affiliations Academic affiliations
  % should list Department, University, City, Region, Country Industry
  % affiliations should list Company, City, Region, Country

  % You can specify symbols, otherwise they are numbered in order. Ideally, you
  % should not use this facility. Affiliations will be numbered in order of
  % appearance and this is the preferred way.
  \icmlsetsymbol{equal}{*}

  \begin{icmlauthorlist}
    \icmlauthor{Zhongpai Gao}{uii}
    % Additional co-authors can be added here if needed:
    % \icmlauthor{Meng Zheng}{uii}
    % \icmlauthor{Benjamin Planche}{uii}
    % \icmlauthor{Anwesa Choudhuri}{uii}
    % \icmlauthor{Terrence Chen}{uii}
    % \icmlauthor{Ziyan Wu}{uii}
  \end{icmlauthorlist}

  \icmlaffiliation{uii}{United Imaging Intelligence, Boston, MA, USA}

  \icmlcorrespondingauthor{Zhongpai Gao}{zhongpai.gao@uii-ai.com}

  % You may provide any keywords that you find helpful for describing your
  % paper; these are used to populate the "keywords" metadata in the PDF but
  % will not be shown in the document
  \icmlkeywords{Gaussian Splatting, Neural Rendering, Novel View Synthesis, 3D Reconstruction}

  \vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column listing the
% affiliations and the copyright notice. The command takes one argument, which
% is text to display at the start of the footnote. The \icmlEqualContribution
% command is standard text for equal contribution. Remove it (just {}) if you
% do not need this facility.

% Use ONE of the following lines. DO NOT remove the command.
% If you have no special notice, KEEP empty braces:
\printAffiliationsAndNotice{}  % no special notice (required even if empty)
% Or, if applicable, use the standard equal contribution text:
% \printAffiliationsAndNotice{\icmlEqualContribution}

\begin{abstract}
  High-dimensional Gaussian Splatting methods like 6DGS and 7DGS achieve impressive view-dependent and dynamic scene modeling through conditional N-dimensional Gaussians. However, their monolithic covariance design derives all conditioning effects from a shared joint covariance matrix, requiring expensive cubic-complexity matrix inversions in every forward pass. We introduce \modelname{}, a direct parameterization that explicitly represents each conditioning component with efficient, independent parameters. Instead of deriving parameters through covariance inversion, we directly parameterize opacity modulation via Cholesky precision factors and position shifts via displacement matrices, eliminating the need for matrix inversions while maintaining numerical stability. Our key contribution is learnable per-dimension coupling parameters that control how strongly position shifts respond to query deviations, enabling adaptive interpolation from no position conditioning to full N-DGS-style coupling. This direct parameterization achieves up to 17\% higher rendering throughput across diverse benchmarks while matching or exceeding N-DGS quality. Our approach offers a principled, efficient alternative to implicit covariance-based conditioning in high-dimensional Gaussian splatting.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
3D Gaussian Splatting (3DGS)~\citep{3dgs} revolutionized novel view synthesis by achieving photorealistic quality at real-time speeds through explicit 3D Gaussian primitives. However, real-world scenes exhibit complex phenomena beyond static representations: specular reflections, temporal dynamics, and view-dependent effects that require richer modeling capacity.

Recent extensions elevate Gaussians into higher-dimensional spaces to capture these phenomena. N-dimensional Gaussians (N-DG)~\citep{diolatzis2024n} pioneered this direction, refined by 6D Gaussian Splatting (6DGS)~\citep{gao20246dgs} which models position and viewing direction jointly for view-dependent effects, and 7D Gaussian Splatting (7DGS)~\citep{gao20257dgs} which adds temporal dimensions for dynamic scenes. Alternative kernel designs like Deformable Beta Splatting (DBS)~\citep{liu2025deformable} improve geometric fidelity but remain limited to spatial dimensions.

Despite their expressive power, these N-dimensional formulations suffer from a fundamental limitation: their \textit{implicit} parameterization. All conditioning effects---position shifts, opacity modulation, and covariance adjustments---are derived from a single joint covariance matrix via conditional Gaussian formulas. Computing any single effect requires inverting the conditioning block with cubic complexity operations per Gaussian, creating computational overhead that accumulates across millions of primitives. Moreover, this implicit derivation offers limited control: position and opacity conditioning are rigidly tied through the shared covariance structure, preventing independent adjustment of their coupling strengths. The monolithic design also risks numerical instabilities when covariance matrices become ill-conditioned.

\textbf{Our Insight.} Conditional Gaussian formulas are elegant mathematically but computationally inefficient. Each conditioning effect can be \textbf{directly parameterized} with efficient, independent parameters instead of being implicitly derived through matrix algebra. Opacity modulation only needs a precision matrix for distance computation---parameterize it directly via Cholesky factors. Position shifts only need a displacement matrix scaled by query deviations---parameterize it directly with bounded weights. By replacing implicit derivation with explicit parameters, we eliminate matrix inversions while gaining fine-grained control over each effect independently.

We introduce \textbf{\modelname{}}, a direct parameterization for efficient conditional Gaussian splatting. The `\textbf{d}' signifies our principle: \textbf{d}irect parameterization---explicit \textbf{d}isplacement matrices for position and \textbf{d}ecay factors for opacity, enabling \textbf{d}ynamic Gaussian behavior. Like N-DGS, we enable Gaussians to adaptively shift position and modulate opacity based on queries, but replace implicit covariance derivation with efficient explicit parameters. Our key contributions are:


\begin{enumerate}[leftmargin=*]
\item \textbf{Direct Opacity Parameterization:} We directly parameterize opacity decay via Cholesky-factored precision matrices, eliminating the need for matrix inversions while guaranteeing numerical stability through direct triangular operations.

\item \textbf{Direct Position Displacement:} Position shifts are directly parameterized via displacement matrices with spatial scaling and bounded activations, ensuring tight displacement bounds without requiring covariance inversion.

\item \textbf{Learnable Coupling Strength:} Per-dimension coupling parameters control how strongly position shifts respond to query deviations, enabling adaptive interpolation from no position conditioning to full coupling through the shared precision matrix.

\item \textbf{Efficient CUDA Implementation:} We implement optimized CUDA kernels for both our method and N-DGS baseline, fusing conditional operations with register-level computation for fair comparison (Note that the original N-DGS only implements in PyTorch).
\end{enumerate}

Experiments demonstrate that \modelname{} matches or exceeds N-DGS quality across all benchmarks while achieving up to 17\% higher rendering throughput.

\section{Related Work}
\label{sec:related_work}

\textbf{3D Gaussian Splatting and Extensions.}
3D Gaussian Splatting (3DGS)~\cite{3dgs} revolutionized novel view synthesis by representing scenes as collections of anisotropic 3D Gaussians, enabling real-time rendering with competitive quality. The method's success has inspired numerous extensions addressing anti-aliasing~\cite{yu2023mipsplattingaliasfree3dgaussian,yan2024multiscale3dgaussiansplatting}, compression~\cite{lee2024compact,morgenstern2024compact}, and semantic applications~\cite{shi2023languageembedded3dgaussians,zhou2024feature3dgssupercharging3d}. However, 3DGS's fixed Gaussian primitives and spherical harmonic color encoding struggle with complex view-dependent effects and dynamic scenes.

\textbf{N-Dimensional Gaussian Methods.}
To address view-dependent effects, N-dimensional Gaussians (N-DG)~\cite{diolatzis2024n} extended Gaussian representations into higher dimensions, refined by 6D Gaussian Splatting (6DGS)~\cite{gao20246dgs} which models position and viewing direction jointly. 7D Gaussian Splatting (7DGS)~\cite{gao20257dgs} further incorporates temporal dimensions for dynamic scenes. These methods achieve impressive results but their monolithic covariance design creates computational overhead and can introduce numerical instabilities when covariance matrices become ill-conditioned.

\textbf{Efficient Parameterizations.}
Various parameterization strategies have been explored for covariance matrices in Gaussian splatting. The original 3DGS uses quaternion-based rotation with diagonal scaling, ensuring positive semi-definiteness while maintaining interpretability. For higher dimensions, methods resort to Cholesky factorization, which loses explicit orthogonality but generalizes to arbitrary dimensions. Recent work on deformable splatting~\cite{liu2025deformable} explores adaptive kernel shapes, though these remain limited to spatial dimensions without addressing the coupling problem in conditional distributions.

\textbf{Dynamic Scene Modeling.}
Early NeRF extensions modeled dynamics through deformation fields: D-NeRF~\cite{pumarola2020dnerfneuralradiancefields}, Nerfies~\cite{park2021nerfies}, and HyperNeRF~\cite{park2021hypernerf}. For Gaussian Splatting, 4D Gaussian Splatting (4DGS)~\cite{yang2023real} extends primitives temporally, while Dynamic 3D Gaussians~\cite{luiten2023dynamic} combines canonical representations with learned deformations. 7DGS unifies spatial, temporal, and angular dimensions but enforces tight coupling through shared covariance structures, preventing independent control of position and opacity modulation.

Our work addresses these limitations by introducing direct parameterization of conditioning effects, eliminating matrix inversions while enabling learnable control over position-opacity coupling strength through per-dimension parameters.

\section{Preliminary}

Our work builds upon the family of Gaussian Splatting methods. To contextualize our contribution, we first detail the progression from static 3D models to the more complex 6D and 7D formulations, highlighting the trade-offs between expressive power and computational cost at each step.

\paragraph{3D Gaussian Splatting}
3D Gaussian Splatting (3DGS) revolutionized novel view synthesis by representing scenes as a collection of anisotropic 3D Gaussians. Each Gaussian is a primitive defined by a set of learnable attributes that describe its geometry and appearance. Its geometry is determined by a mean vector \(\boldsymbol{\mu} \in \mathbb{R}^3\) specifying its center in 3D space, and a covariance matrix \(\boldsymbol{\Sigma} \in \mathbb{R}^{3 \times 3}\) defining its shape, scale, and orientation. To ensure physical plausibility and simplify optimization, \(\boldsymbol{\Sigma}\) is parameterized by a diagonal scaling matrix \(\mathbf{S} = \operatorname{diag}(s_x, s_y, s_z)\) and a rotation matrix \(\mathbf{R}\) (represented as a quaternion), such that \(\boldsymbol{\Sigma} = \mathbf{R} \mathbf{S} \mathbf{S}^\top \mathbf{R}^\top\). The appearance of each Gaussian is governed by a scalar opacity \(\alpha \in [0, 1]\) that controls its transparency, and a view-dependent color modeled using Spherical Harmonics (SH). The color \(c\) for a given viewing direction \(d\) is computed as a sum over the SH basis functions \(Y_{\ell m}\):
\begin{equation}
c(d) = \sum_{\ell=0}^{N} \sum_{m=-\ell}^{\ell} \beta_{\ell m} Y_{\ell m}(d),
\end{equation}
where \(\beta_{\ell m}\) are the learnable SH coefficients. 3DGS achieves state-of-the-art quality and real-time rendering for static scenes. However, its representation is fundamentally static and cannot model geometric changes arising from complex view-dependent effects or temporal dynamics.

\paragraph{N-Dimensional Gaussian Splatting}

To capture view-dependent and temporal effects, recent methods extend 3DGS into higher dimensions: 6DGS for view-dependent phenomena and 7DGS for dynamic scenes. Each primitive is an N-dimensional Gaussian over position $\mathbf{X}_p \in \mathbb{R}^3$ and query variables $\mathbf{X}_q \in \mathbb{R}^C$ (view direction for 6DGS with $C=3$, view+time for 7DGS with $C=4$):
\begin{align}
\mathbf{X} = \begin{pmatrix} \mathbf{X}_p \\ \mathbf{X}_q \end{pmatrix} \sim \mathcal{N}\!\left(
\begin{pmatrix}
\boldsymbol{\mu}_p \\ \boldsymbol{\mu}_q
\end{pmatrix},
\begin{pmatrix}
\boldsymbol{\Sigma}_{pp} & \boldsymbol{\Sigma}_{pq} \\
\boldsymbol{\Sigma}_{pq}^\top & \boldsymbol{\Sigma}_{qq}
\end{pmatrix}
\right).
\end{align}
At query $\mathbf{q}$, conditioning yields a 3D Gaussian with parameters:
\begin{align}
\alpha_{\text{cond}} &= \alpha \cdot \exp\left(-\lambda\,(\mathbf{q}-\boldsymbol{\mu}_q)^\top \boldsymbol{\Sigma}_{qq}^{-1}(\mathbf{q}-\boldsymbol{\mu}_q)\right),\\
\boldsymbol{\mu}_{\text{cond}} &= \boldsymbol{\mu}_p + \boldsymbol{\Sigma}_{pq}\,\boldsymbol{\Sigma}_{qq}^{-1}(\mathbf{q}-\boldsymbol{\mu}_q), \\
\boldsymbol{\Sigma}_{\text{cond}} &= \boldsymbol{\Sigma}_{pp} - \boldsymbol{\Sigma}_{pq}\,\boldsymbol{\Sigma}_{qq}^{-1}\,\boldsymbol{\Sigma}_{pq}^\top,
\end{align}
where $\lambda$ is a scaling factor (default 0.35). This formulation requires inverting $\boldsymbol{\Sigma}_{qq}$ ($O(C^3)$ operations) for every Gaussian in every forward pass, causing computational overhead and potential numerical instability during training. This motivates our simpler parameterization.

\section{Our Approach}

N-DGS methods derive all conditioning effects from a single $(3+C) \times (3+C)$ joint covariance matrix via conditional Gaussian formulas. To compute any effect (position shift, opacity modulation, or covariance adjustment), they must invert the $C \times C$ conditioning block $\boldsymbol{\Sigma}_{qq}$, requiring $O(C^3)$ operations per Gaussian per forward pass.

We propose \modelname{}, which replaces this \textbf{implicit derivation} with \textbf{direct parameterization}. Rather than computing conditioning effects through matrix algebra, we explicitly parameterize each component with dedicated, efficient parameters. This direct approach eliminates cubic-complexity inversions while providing independent control over each conditioning effect.

\subsection{Overview}

Our approach directly parameterizes conditional Gaussian slicing with two explicit components. Given the deviation $\boldsymbol{\delta} = q - \mu_q$ from the canonical view/time:
\begin{align}
\alpha_{\text{cond}} &= \alpha \cdot \exp\left(-\|\mathbf{L}^\top \boldsymbol{\delta}\|_2^2\right), & &\text{(Opacity)} \\
\mu_{\text{cond}} &= \mu_p + \mathbf{V}_{pq} \cdot \tilde{\mathbf{V}}_{qq} \cdot \boldsymbol{\delta}, & &\text{(Position)} \end{align}
where $\mathbf{L} \in \mathbb{R}^{C \times C}$ is the directly parameterized Cholesky factor, $\mathbf{V}_{qq} = \mathbf{L}\mathbf{L}^\top$ is the precision matrix computed directly from $\mathbf{L}$ without inversion, $\mathbf{V}_{pq} = \bar{s}\boldsymbol{\Theta}_{pq} \in \mathbb{R}^{3 \times C}$ is the displacement matrix (mean spatial scale $\bar{s}$ times normalized $\boldsymbol{\Theta}_{pq}$ with $\|\boldsymbol{\Theta}_{pq}\|_F = 1$), and $\tilde{\mathbf{V}}_{qq} = \text{diag}(\boldsymbol{\Lambda}) \mathbf{V}_{qq}$ is the scaled precision with learnable per-dimension coupling strengths $\boldsymbol{\Lambda} \in [0,1]^C$ (dimensions of the same type share parameters). The spatial covariance $\boldsymbol{\Sigma}_{pp}$ remains independent of the query, using the standard 3DGS quaternion-scale parameterization ($\boldsymbol{\Sigma}_{pp} = \mathbf{R}\mathbf{S}\mathbf{S}^\top \mathbf{R}^\top$) for seamless compatibility with existing pipelines, including densification strategies.

For 7DGS ($C=4$), we further \textbf{separate view and time} by using a block-diagonal precision matrix with zero cross-terms:
\begin{equation}
\mathbf{V}_{qq} = \mathbf{L}\mathbf{L}^\top = \begin{pmatrix}
\mathbf{V}_{vv} & \mathbf{0} \\
\mathbf{0} & v_t
\end{pmatrix},
\end{equation}
where $\mathbf{V}_{vv} \in \mathbb{R}^{3 \times 3}$ is the view precision and $v_t \in \mathbb{R}$ is the time precision. This separation eliminates coupling between spatial view directions and temporal dynamics, reflecting the physical intuition that view-dependent effects and temporal changes are independent phenomena.


\begin{table}[t!]
\centering
\caption{Comparison between N-DGS (implicit derivation) and \modelname{} (direct parameterization). Our approach directly parameterizes opacity and position with dedicated parameters, eliminating matrix inversions. The coupling strength $\Lambda$ controls position response to queries.}
\label{tab:comparison_overview}
\setlength{\tabcolsep}{4pt}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Aspect} & \textbf{N-DGS (implicit)} & \textbf{\modelname{} (direct)} & \textbf{Benefit} \\
\midrule
Opacity & $\alpha \cdot \exp(-\lambda\boldsymbol{\delta}^\top\boldsymbol{\Sigma}_{qq}^{-1}\boldsymbol{\delta})$ & $\alpha \cdot \exp(-\|\mathbf{L}^\top\boldsymbol{\delta}\|^2)$ & No inversion \\
Position & $\boldsymbol{\mu}_p + \boldsymbol{\Sigma}_{pq}\boldsymbol{\Sigma}_{qq}^{-1}\boldsymbol{\delta}$ & $\boldsymbol{\mu}_p + \mathbf{V}_{pq}\tilde{\mathbf{V}}_{qq}\boldsymbol{\delta}$ & Direct $\mathbf{V}_{pq}$ \\
Covariance & $\boldsymbol{\Sigma}_{pp} - \boldsymbol{\Sigma}_{pq}\boldsymbol{\Sigma}_{qq}^{-1}\boldsymbol{\Sigma}_{pq}^\top$ & $\boldsymbol{\Sigma}_{pp}$ (direct) & No correction term \\
\bottomrule
\end{tabular}%
}
\end{table}

Table~\ref{tab:comparison_overview} compares our direct parameterization with N-DGS's implicit derivation. The key innovation is \textbf{learnable coupling strength} via per-dimension parameters $\Lambda$: when $\Lambda = \mathbf{0}$, position becomes query-independent ($\boldsymbol{\mu}_{\text{cond}} = \boldsymbol{\mu}_p$); when $\Lambda = \mathbf{1}$, position couples through the shared precision matrix $\mathbf{V}_{qq}$ like N-DGS. For the simplified \textbf{\textit{d}GS-O} variant, we disable position conditioning entirely ($\boldsymbol{\mu}_{\text{cond}} = \boldsymbol{\mu}_p$) and use only direct opacity parameterization for maximum efficiency. The following subsections detail the direct parameterization of each component.

\subsection{Opacity Conditioning}

In N-DGS, opacity is modulated as $\alpha_{\text{cond}} = \alpha \cdot \exp(-\lambda\,(\mathbf{q}-\boldsymbol{\mu}_q)^\top \boldsymbol{\Sigma}_{qq}^{-1}(\mathbf{q}-\boldsymbol{\mu}_q))$, requiring matrix inversion.

Our formulation directly parameterizes the precision matrix via its Cholesky factor $\mathbf{L}$, computing $\mathbf{L}^\top \boldsymbol{\delta}$ followed by $\|\cdot\|_2^2$ through a triangular matrix-vector multiplication that is highly parallelizable on GPUs. This eliminates the need for matrix inversion. Parameterizing the Cholesky factor directly with exponential activation on the diagonal guarantees that $\mathbf{V}_{qq} = \mathbf{L}\mathbf{L}^\top$ is always symmetric positive definite, avoiding numerical instabilities from ill-conditioned matrices.

For 7DGS, the block-diagonal structure enables \textbf{separated view and time opacity}. With $\boldsymbol{\delta} = (\boldsymbol{\delta}_v, \delta_t)^\top$ where $\boldsymbol{\delta}_v \in \mathbb{R}^3$ is the view deviation and $\delta_t \in \mathbb{R}$ is the time deviation, the opacity factors as:
\begin{equation}
\alpha_{\text{cond}} = \alpha \cdot \exp\left(-\left(\boldsymbol{\delta}_v^\top \mathbf{V}_{vv} \boldsymbol{\delta}_v + \delta_t^2 v_t\right)\right).
\end{equation}
This factorization allows view-dependent and time-dependent opacity to be controlled independently, with view precision $\mathbf{V}_{vv}$ parameterized by a $3 \times 3$ Cholesky factor (6 parameters) and time precision $v_t$ by a single scalar, for a total of 7 parameters instead of 10 for a full $4 \times 4$ precision matrix.

\subsection{Position Conditioning}
\label{subsec:position}

In N-DGS, position is computed as $\boldsymbol{\mu}_{\text{cond}} = \boldsymbol{\mu}_p + \boldsymbol{\Sigma}_{pq}\boldsymbol{\Sigma}_{qq}^{-1}(\mathbf{q}-\boldsymbol{\mu}_q)$, where both $\boldsymbol{\Sigma}_{pq}$ and $\boldsymbol{\Sigma}_{qq}^{-1}$ are implicitly derived from the joint covariance matrix through Cholesky decomposition and matrix inversion.

We replace this implicit derivation with \textbf{direct parameterization} of position shifts, combined with learnable coupling strength control:
\begin{equation}
\boldsymbol{\mu}_{\text{cond}} = \boldsymbol{\mu}_p + \underbrace{\bar{s} \boldsymbol{\Theta}_{pq}}_{\mathbf{V}_{pq}} \cdot \underbrace{\text{diag}(\Lambda) \mathbf{V}_{qq}}_{\tilde{\mathbf{V}}_{qq}} \cdot \boldsymbol{\delta},
\end{equation}
where $\bar{s} = \text{mean}(s_x, s_y, s_z)$ is the spatial scale, $\boldsymbol{\Theta}_{pq} \in \mathbb{R}^{3 \times C}$ are learnable parameters normalized to unit Frobenius norm ($\|\boldsymbol{\Theta}_{pq}\|_F = 1$), $\text{diag}(\boldsymbol{\Lambda}) \in \mathbb{R}^{C \times C}$ contains learnable per-dimension coupling strengths in $[0,1]$, and $\mathbf{V}_{qq} = \mathbf{L}\mathbf{L}^\top$ is the precision matrix shared with opacity conditioning. The coupling strength $\boldsymbol{\Lambda}$ controls how strongly position shifts respond to query deviations:
\begin{itemize}[leftmargin=*]
\item When $\Lambda = \mathbf{0}$: $\tilde{\mathbf{V}}_{qq} = \mathbf{0}$, position becomes query-independent ($\boldsymbol{\mu}_{\text{cond}} = \boldsymbol{\mu}_p$, no conditioning).
\item When $\Lambda = \mathbf{1}$: $\tilde{\mathbf{V}}_{qq} = \mathbf{V}_{qq}$, position and opacity couple through the shared precision matrix $\mathbf{V}_{qq}$ (similar to N-DGS, but with direct $\mathbf{V}_{pq}$ instead of derived $\boldsymbol{\Sigma}_{pq}$).
\item Intermediate values: Partial coupling with per-dimension control. Different query dimensions (e.g., view vs. time) can have different coupling strengths, with dimensions of the same type (e.g., the three view dimensions) sharing parameters.
\end{itemize}
This direct parameterization eliminates covariance inversion while providing fine-grained control over position conditioning strength. Each Gaussian learns optimal per-dimension coupling during training.

Inspired by N-DGS's design principle that displacement should scale with spatial extent (so small Gaussians produce small displacements), our displacement matrix $\mathbf{V}_{pq} = \bar{s} \boldsymbol{\Theta}_{pq}$ adopts this spatial scaling idea but goes beyond their implicit Cholesky-based approach through explicit parameterization. By normalizing $\boldsymbol{\Theta}_{pq}$ to unit Frobenius norm ($\|\boldsymbol{\Theta}_{pq}\|_F = 1$), displacements scale proportionally with the average Gaussian extent $\bar{s} = \text{mean}(s_x, s_y, s_z)$, with each row uniformly bounded by $\|\mathbf{V}_{pq}^{(i)}\|_2 \leq \bar{s}$. This preserves the inductive bias coupling displacement scale to Gaussian size, while enabling tighter bounds, uniform treatment across spatial dimensions, and better handling of anisotropic Gaussians. (See Appendix~\ref{sec:spatial_scaling_proof} for detailed comparison showing how our direct parameterization realizes this principle more effectively than N-DGS.)


\subsection{CUDA Implementation}

Since the original N-DGS methods (6DGS and 7DGS) use PyTorch's autograd for conditional Gaussian operations without custom CUDA kernels, we implement both our disentangled formulation and the N-DGS baseline with optimized CUDA kernels for fair comparison. Our implementation uses a custom kernel that fuses conditional position and opacity computations in a single pass, as shown in Algorithm~\ref{alg:slice}.

\textbf{CUDA Kernel Design.} The kernel is templated on dimension $C$ (3 for 6DGS, 4 for 7DGS) with compile-time loop unrolling for optimal performance. Each thread processes one Gaussian primitive independently, enabling perfect parallelization across millions of Gaussians. For opacity computation (Line 2 in Algorithm~\ref{alg:slice}), we compute the triangular matrix-vector product $\mathbf{L}^\top \boldsymbol{\delta}$ directly in registers using the lower-triangular structure—only $C(C+1)/2$ multiplications are needed (6 for 6DGS, 10 for 7DGS), followed by a dot product for $\|\cdot\|_2^2$. For position displacement (Lines 3-5), we construct $\mathbf{V}_{pq}$ via diagonal scaling of $\tanh$ activations, then compute the scaled precision matrix $\tilde{\mathbf{V}}_{qq}$ by applying per-dimension scaling $\text{diag}(\Lambda)$ to $\mathbf{L}\mathbf{L}^\top$. All matrix operations use registers rather than shared memory, as the small matrix sizes fit entirely in per-thread registers. This fused design minimizes global memory bandwidth by avoiding intermediate tensor allocations and maximizes GPU occupancy through register-level computation.

\textbf{Modular Compatibility.} Our implementation provides compile-time options to enable different levels of conditioning: (1) standard 3DGS with no conditioning, (2) \textit{d}GS-O with opacity modulation only, and (3) full \textit{d}GS with both opacity and position modulation. This modularity maintains compatibility with existing 3DGS codebases while allowing practitioners to select the appropriate complexity-performance tradeoff for their application.


\begin{algorithm}[tb]
\caption{\textit{d}GS Conditional Gaussian Slicing}
\label{alg:slice}
\begin{algorithmic}[1]
\REQUIRE Gaussian parameters: $\mu_p$, $\mu_q$, $\alpha$, $\mathbf{S}$, $\boldsymbol{\Theta}_{pq}$, $\mathbf{L}$, $\Lambda$
\REQUIRE Query point: $q$ (view/time)
\STATE $\boldsymbol{\delta} \gets q - \mu_q$ %\Comment{Compute query deviation}
\STATE $\alpha_{\text{cond}} \gets \alpha \cdot \exp(-\|\mathbf{L}^\top \boldsymbol{\delta}\|^2)$ % \Comment{Conditional opacity}
\IF{\textit{d}GS-O variant}
    \STATE \return  $\alpha_{\text{cond}}, \mu_p$ \Comment{Early exit for opacity only}
\ENDIF
\STATE $\mathbf{V}_{pq} \gets \text{mean}(\text{diag}(\mathbf{S})) \cdot \boldsymbol{\Theta}_{pq}$ % \Comment{Construct displacement matrix}
\STATE $\tilde{\mathbf{V}}_{qq} \gets \text{diag}(\boldsymbol{\Lambda}) \mathbf{L}\mathbf{L}^\top$ % \Comment{Scaled precision for coupling}
\STATE $\mu_{\text{cond}} \gets \mu_p + \mathbf{V}_{pq} \cdot \tilde{\mathbf{V}}_{qq} \cdot \boldsymbol{\delta}$ % \Comment{Conditional position}
\STATE \return $\alpha_{\text{cond}}$, $\mu_{\text{cond}}$ 

\end{algorithmic}
\end{algorithm}

\section{Experiments}

\subsection{Experimental Setup}

\textbf{Datasets.}
We evaluate \modelname{} across five diverse datasets covering both static and dynamic scenes:
\begin{itemize}[leftmargin=*]
    \item \textit{Static scenes:} (1) \textit{NeRF Synthetic}~\cite{mildenhall2021nerf} with 8 synthetic scenes under controlled lighting for geometric reconstruction assessment; (2) \textit{Mip-NeRF 360}~\cite{barron2021mipnerf} with 9 unbounded real-world scenes challenging view synthesis; and (3) \textit{6DGS PBR}~\cite{gao20246dgs} featuring 7 physically-based scenes with complex view-dependent effects.

    \item \textit{Dynamic scenes:} (4) \textit{D-NeRF}~\cite{pumarola2020dnerfneuralradiancefields} with 8 synthetic sequences featuring various deformation types; and (5) \textit{7DGS-PBR}~\cite{gao20257dgs} containing 6 physically-based dynamic scenes with complex spatio-temporal-angular correlations.
\end{itemize}

\textbf{Evaluation Metrics.}
We assess reconstruction quality using Peak Signal-to-Noise Ratio (PSNR), Structural Similarity Index (SSIM)~\cite{wang2004image}, and Learned Perceptual Image Patch Similarity (LPIPS)~\cite{zhang2018perceptual}. We also report rendering speed (FPS), training time (minutes), and Gaussian count (K) to evaluate computational efficiency.

\textbf{Implementation Details.}
We compare three configurations: (1) \textbf{N-DGS}: full N-DGS formulation with joint position and opacity conditioning (entangled baseline), (2) \textbf{\textit{d}GS-O}: our disentangled approach with opacity conditioning only ($\mu_{\text{cond}} = \mu_p$, no position displacement), and (3) \textbf{\textit{d}GS}: our full disentangled approach with both opacity conditioning and learnable per-dimension position coupling via $\Lambda$. Training uses Adam optimizer for 30K iterations. Proper initialization is critical for stable training: we initialize the view mean $\mu_d$ as random unit vectors (with time in $[0, 1]$ for 7DGS), the displacement parameters $\boldsymbol{\Theta}_{pq}$ from $\mathcal{N}(0, 0.01^2)$ to start with small displacements, and the Cholesky factor $\mathbf{L}$ with diagonal elements $\log(2.0)$ and off-diagonal elements 0.

\begin{table*}[t]
\centering
\caption{Comparison of methods on Mip-NeRF 360, NeRF Synthetic, and 6DGS PBR datasets with MCMC densification. Mean values across all scenes are reported. Number of Gaussians (K) and train time (min). FPS and FPS$_{\text{CUDA}}$ represent that conditional Gaussian operations are implemented in PyTorch and CUDA, respectively. Note that, N-DGS, \textit{d}GS-O, and \textit{d}GS are integrated with TCGS to speed up rendering.}
\label{tab:mcmc_comparison}
\resizebox{0.85\textwidth}{!}{%
\begin{tabular}{@{}cl|rrrr|ccc@{}}
\toprule
Dataset & Method & \#Gauss $\downarrow$ & Train $\downarrow$ & FPS $\uparrow$& FPS$_{\text{CUDA}}$ $\uparrow$& PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\multirow{4}{*}{\makecell{\texttt{Mip-NeRF}\\\texttt{360}}}
& 3DGS & 3,104 & \bf 40.16 & 47.62 & --- & 27.06 & 0.788 & 0.237 \\
& N-DGS & 2,852 & 51.23 & 33.79 & 75.31 & 26.88 & 0.775 & 0.263 \\
& \textit{d}GS-O (\textbf{Ours}) & 2,852 & 48.11 & \bf 91.43 & 200.17 & 27.36 & 0.788 & 0.242 \\
& \textit{d}GS (\textbf{Ours}) & 2,574 & 49.85 & 43.72 & \bf 200.56 & \bf 27.67 & \bf 0.794 & \bf 0.227 \\
\midrule
\multirow{4}{*}{\makecell{\texttt{NeRF}\\\texttt{Synthetic}}}
& 3DGS & 300 & \bf 7.80 & 303.09 & --- & 33.72 & 0.971 & 0.028 \\
& N-DGS & 300 & 10.43 & 179.12 & 348.95 & 34.04 & \bf 0.973 & 0.027 \\
& \textit{d}GS-O (\textbf{Ours}) & 300 & 9.71 & \bf 400.96 & 587.84 & 34.25 & \bf 0.973 & \bf 0.026 \\
& \textit{d}GS (\textbf{Ours}) & 300 & 10.26 & 267.33 & \bf 593.94 & \bf 34.34 & \bf 0.973 & \bf 0.026 \\
\midrule
\multirow{4}{*}{\makecell{\texttt{6DGS}\\\texttt{PBR}}}
& 3DGS & 163 & 53.28& 183.78 & --- & 30.89 & 0.938 & 0.077 \\
& N-DGS & 163 & 54.80 & 292.60 & 476.79 & 40.06 & \bf 0.976 & 0.039 \\
& \textit{d}GS-O (\textbf{Ours}) & 163 & 48.72  & \bf 467.06 & \bf 588.23& 39.51 & 0.971 & 0.046 \\
& \textit{d}GS (\textbf{Ours}) & 163 & \bf 47.54 & 364.74 & 551.02 & \bf 40.34 & \bf 0.976 & \bf 0.038 \\
\bottomrule
\end{tabular}}
\end{table*}


% Merged Summary Results Table
\begin{table*}[h]
\centering
\caption{Mean results across all datasets: 3 static (NeRF Synthetic, Mip-NeRF 360, 6DGS PBR) and 2 dynamic (D-NeRF, 7DGS PBR). Top: quality metrics (PSNR, SSIM, LPIPS). Bottom: efficiency metrics (\#Gaussians in K, training time in mins, rendering FPS). See Appendix for per-scene details.}
\label{tab:summary_results}
\resizebox{0.85\textwidth}{!}{%
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
 & \multicolumn{3}{c|}{N-DGS} & \multicolumn{3}{c|}{\textit{d}GS-O (\textbf{Ours})} & \multicolumn{3}{c}{\textit{d}GS (\textbf{Ours})} \\
\midrule
\multicolumn{10}{c}{\textbf{Quality Metrics}} \\
\midrule
Dataset & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\texttt{NeRF Synthetic} & 33.38 & 0.970 & 0.032 & 33.84 & 0.971 & 0.030 & \textbf{33.91} & \textbf{0.972} & \textbf{0.029} \\
\texttt{Mip-NeRF 360} & 26.40 & 0.739 & 0.316 & 27.18 & 0.779 & 0.252 & \textbf{27.37} & \textbf{0.781} & \textbf{0.251} \\
\texttt{6DGS PBR} & 37.11 & 0.971 & 0.054 & 36.54 & 0.965 & 0.059 & \textbf{37.75} & \textbf{0.972} & \textbf{0.052} \\
\midrule
\texttt{D-NeRF} & 33.13 & 0.968 & 0.038 & 30.92 & 0.959 & 0.044 & \textbf{33.59} & \textbf{0.971} & \textbf{0.032} \\
\texttt{7DGS PBR} & 31.75 & 0.955 & 0.055 & 31.70 & 0.955 & 0.053 & \textbf{32.18} & \textbf{0.956} & \textbf{0.050} \\
\midrule
\midrule
\multicolumn{10}{c}{\textbf{Efficiency Metrics}} \\
\midrule
Dataset & \#Gauss $\downarrow$ & Time $\downarrow$ & FPS$_{\text{CUDA}}$ $\uparrow$ & \#Gauss $\downarrow$ & Time $\downarrow$ & FPS$_{\text{CUDA}}$ $\uparrow$ & \#Gauss $\downarrow$ & Time $\downarrow$ & FPS$_{\text{CUDA}}$ $\uparrow$ \\
\midrule
\texttt{NeRF Synthetic} & \textbf{171} & 6.35 & 615.57 & 201 & \textbf{5.83} & \textbf{754.98} & 201 & 6.84 & 704.84 \\
\texttt{Mip-NeRF 360} & \textbf{1273} & 27.09 & 147.46 & 2363 & \textbf{25.58} & \textbf{279.30} & 2463 & 28.21 & 237.82 \\
\texttt{6DGS PBR} & \textbf{84} & \textbf{26.77} & 678.40 & 118 & 30.72 & \textbf{696.31} & 134 & 33.80 & 653.44 \\
\midrule
\texttt{D-NeRF} & \textbf{69} & 14.08 & 852.01 & 70 & \textbf{13.20} & \textbf{948.77} & 85 & 15.56 & 818.41 \\
\texttt{7DGS PBR} & \textbf{162} & 38.14 & 709.83 & 227 & \textbf{34.58} & \textbf{835.24} & 226 & 36.96 & 753.77 \\
\bottomrule
\end{tabular}%
}
\end{table*}


\subsection{Results and Analysis}

The comparison between our three configurations reveals distinct efficiency-quality trade-offs. \textbf{\textit{d}GS-O} uses only opacity conditioning with $\mu_{\text{cond}} = \mu_p$ (no position displacement). On static scenes (NeRF Synthetic, Mip-NeRF 360, 6DGS PBR), it achieves strong quality: +0.46 dB on NeRF Synthetic, +0.78 dB on Mip-NeRF 360, but -0.57 dB on 6DGS PBR compared to N-DGS. The efficiency gains are substantial: 23\% faster rendering (754.98 vs 615.57 FPS on NeRF Synthetic), 89\% faster on Mip-NeRF 360 (279.30 vs 147.46 FPS). Notably, when primitive count is large (e.g., \texttt{Mip-NeRF 360}), \textit{d}GS-O achieves 89\% faster rendering despite using 86\% more primitives (2363K vs 1273K), demonstrating that computational efficiency is dominated by per-primitive operations rather than primitive count. However, on dynamic scenes, the lack of position displacement significantly impacts quality: -2.21 dB on D-NeRF, indicating that temporal dynamics benefit more from position conditioning than static view-dependent effects.

\textbf{\textit{d}GS} adds learnable per-dimension coupling strength via $\Lambda$, enabling each Gaussian to adaptively control how strongly position shifts respond to query deviations. When $\Lambda = \mathbf{0}$, position becomes query-independent (no conditioning); when $\Lambda = \mathbf{1}$, position couples through the shared precision matrix $\mathbf{V}_{qq}^{-1}$ like N-DGS. This recovers quality losses: on D-NeRF it achieves +0.39 dB over N-DGS (33.52 vs 33.13 dB), demonstrating that learnable coupling strength control can even surpass N-DGS's fixed coupling. On static scenes, \textit{d}GS consistently matches or exceeds N-DGS (+0.46 dB on NeRF Synthetic, +0.93 dB on Mip-NeRF 360, +0.66 dB on 6DGS PBR), while maintaining competitive rendering speeds (687.87 FPS on NeRF Synthetic, 12\% faster than N-DGS).

For 7DGS PBR, the results are highly competitive: \textit{d}GS-O achieves 31.70 dB (only -0.05 dB vs N-DGS at 31.75 dB) while \textit{d}GS matches this at 31.72 dB (-0.03 dB). Both variants deliver substantial efficiency gains, with \textit{d}GS-O achieving 18\% faster rendering (835.24 vs 709.83 FPS) and 9\% faster training (34.58 vs 38.14 min). The minimal quality gap demonstrates that direct parameterization with learnable coupling strength effectively adapts to scenes with complex spatio-temporal-angular correlations while maintaining efficiency benefits.

\subsection{Ablation Study}

To understand the contribution of each component in our approach, we conduct an ablation study comparing three configurations: (1) \textbf{\textit{d}GS-O}: opacity conditioning only without position displacement, (2) \textbf{\textit{d}GS w/o $\bar{s}$}: position displacement without spatial scaling, and (3) \textbf{\textit{d}GS}: our full method with spatial scaling. Table~\ref{tab:ablation} shows results on NeRF Synthetic and 6DGS PBR datasets. Spatial scaling is critical for quality: adding spatial scaling to position displacement improves PSNR by 2.58 dB on NeRF Synthetic (33.84 vs 31.26) and 2.99 dB on 6DGS PBR (37.77 vs 34.78). Notably, position displacement without proper scaling (dGS w/o $\bar{s}$) performs worse than using no position displacement at all (dGS-O), with drops of -2.58 dB on NeRF Synthetic and -1.76 dB on 6DGS PBR. This validates our design choice to scale displacement by Gaussian size in the displacement parameterization.

\begin{table}[h]
\centering
\caption{Ablation study on NeRF Synthetic and 6DGS PBR datasets. \textit{d}GS w/o $\bar{s}$ removes spatial scaling while \textit{d}GS uses spatial scaling.}
\label{tab:ablation}
\setlength{\tabcolsep}{4pt}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{l|ccc|ccc}
\toprule
 & \multicolumn{3}{c|}{\texttt{NeRF Synthetic}} & \multicolumn{3}{c}{\texttt{6DGS PBR}} \\
Config & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ & PSNR $\uparrow$ & SSIM $\uparrow$ & LPIPS $\downarrow$ \\
\midrule
\textit{d}GS-O & 33.84 & 0.971 & 0.030 & 36.54 & 0.965 & 0.059 \\
\textit{d}GS w/o $\bar{s}$ & 31.26 & 0.957 & 0.046 & 34.78 & 0.951 & 0.075 \\
\textit{d}GS (full) & \textbf{33.91} & \textbf{0.972} & \textbf{0.029} & \textbf{37.75} & \textbf{0.972} & \textbf{0.052} \\
\bottomrule
\end{tabular}}%
\end{table}

\section{Future Work}

The direct parameterization framework naturally extends to additional conditioning effects, analogous to factorized attention mechanisms where opacity and position are \textbf{values} independently modulated by view and time as \textbf{queries}. We can extend both the \textit{value space} (what gets conditioned) and the \textit{query space} (what drives conditioning). For values, we can add \textbf{rotation conditioning} (for orientation-dependent effects) or \textbf{color modulation} (for appearance variations beyond spherical harmonics). For queries, we can incorporate \textbf{lighting direction} (for relighting) or \textbf{material properties} (for physically-based rendering). Each new conditioning effect can be directly parameterized with dedicated parameters and its own learnable coupling strength, avoiding the combinatorial complexity of full joint covariance matrices. For instance, rotation conditioning can directly parameterize orientation changes via quaternion spherical linear interpolation, where a learnable rotation offset is blended with the base rotation based on a coupling strength parameter. This enables view/time-dependent orientation as an independent module without requiring joint covariance computation.

\section{Conclusion}

We presented \modelname{}, a direct parameterization for efficient N-dimensional Gaussian splatting. Our key insight is that conditional Gaussian formulas, while mathematically elegant, are computationally inefficient---each conditioning effect can be directly parameterized with dedicated parameters instead of being implicitly derived through matrix algebra. This direct approach delivers efficiency by eliminating cubic-complexity matrix inversions, stability through explicit Cholesky parameterization, and flexibility via learnable per-dimension coupling strengths that control how strongly each effect responds to queries. Experiments demonstrate up to 17\% higher rendering throughput while matching or exceeding N-DGS quality across diverse benchmarks. The direct parameterization framework naturally extends to additional conditioning effects like rotation, lighting, and materials, providing a principled foundation for future high-dimensional Gaussian representations.

% References section
\bibliographystyle{icml2026}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn
\section{Comparison to N-DGS Displacement Parameterization}
\label{sec:spatial_scaling_proof}

Both N-DGS and \modelname{} scale displacement with spatial scale $s_i$, but differ in how tightly this coupling is achieved and in their parameterization approach.

\textbf{N-DGS: Implicit coupling via Cholesky structure.} N-DGS parameterizes the full $(3+C) \times (3+C)$ covariance via Cholesky decomposition $\boldsymbol{\Sigma} = \mathbf{L}\mathbf{L}^\top$, where $\mathbf{L}$ is lower triangular with diagonal elements $L_{ii} = s_i$ (spatial and query scales) and off-diagonal elements $L_{ij} = \ell_{ij}$ for $i > j$, where $\ell_{ij} \in [-1,1]$ are bounded correlation parameters (initialized to zero).

To illustrate this structure, consider 6DGS with 3 spatial and 3 view dimensions ($C=3$). The Cholesky factor $\mathbf{L} \in \mathbb{R}^{6 \times 6}$ has the form:
\begin{equation}
\mathbf{L} = \left(\begin{array}{rrrrrr}
       s_1 &        0 &        0 &        0 &        0 &        0 \\
\ell_{21} &       s_2 &        0 &        0 &        0 &        0 \\
\ell_{31} & \ell_{32} &       s_3 &        0 &        0 &        0 \\
\ell_{41} & \ell_{42} & \ell_{43} &       s_4 &        0 &        0 \\
\ell_{51} & \ell_{52} & \ell_{53} & \ell_{54} &       s_5 &        0 \\
\ell_{61} & \ell_{62} & \ell_{63} & \ell_{64} & \ell_{65} &       s_6
\end{array}\right),
\end{equation}
where $s_1, s_2, s_3$ are spatial scales, $s_4, s_5, s_6$ are query dimension scales, and $\ell_{ij} \in [-1,1]$ are bounded correlations.

% \resizebox{\textwidth}{!}{%
% \begin{equation}    
% \boldsymbol{\Sigma} = \left(\begin{array}{rrrrrr}
% s_1^2 
% &  
% &  
% &  
% &  
% &  \\
% s_1 \ell_{21} 
% & \ell_{21}^2 + s_2^2 
% &  
% &  
% &  
% &  \\
% s_1 \ell_{31} 
% & \ell_{21}\ell_{31} + s_2\ell_{32} 
% & \ell_{31}^2 + \ell_{32}^2 + s_3^2 
% &  
% &  
% &  \\
% s_1 \ell_{41} 
% & \ell_{21}\ell_{41} + s_2\ell_{42} 
% & \ell_{31}\ell_{41} + \ell_{32}\ell_{42} + s_3\ell_{43} 
% & \ell_{41}^2 + \ell_{42}^2 + \ell_{43}^2 + s_4^2 
% &  
% &  \\
% s_1 \ell_{51} 
% & \ell_{21}\ell_{51} + s_2\ell_{52} 
% & \ell_{31}\ell_{51} + \ell_{32}\ell_{52} + s_3\ell_{53} 
% & \ell_{41}\ell_{51} + \ell_{42}\ell_{52} + \ell_{43}\ell_{53} + s_4\ell_{54} 
% & \ell_{51}^2 + \ell_{52}^2 + \ell_{53}^2 + \ell_{54}^2 + s_5^2 
% &  \\
% s_1 \ell_{61} 
% & \ell_{21}\ell_{61} + s_2\ell_{62} 
% & \ell_{31}\ell_{61} + \ell_{32}\ell_{62} + s_3\ell_{63} 
% & \ell_{41}\ell_{61} + \ell_{42}\ell_{62} + \ell_{43}\ell_{63} + s_4\ell_{64} 
% & \ell_{51}\ell_{61} + \ell_{52}\ell_{62} + \ell_{53}\ell_{63} + \ell_{54}\ell_{64} + s_5\ell_{65} 
% & \ell_{61}^2 + \ell_{62}^2 + \ell_{63}^2 + \ell_{64}^2 + \ell_{65}^2 + s_6^2
% \end{array}\right)
% \end{equation}}

The displacement block $\boldsymbol{\Sigma}_{pq} = (\mathbf{L}\mathbf{L}^\top)_{1:3,4:6}$ emerges from matrix multiplication. Computing via row-column dot products yields:
\begin{equation}
\boldsymbol{\Sigma}_{pq} = \left(\begin{array}{rrr}
                           s_1 \ell_{41} &                            s_1 \ell_{51} &                            s_1 \ell_{61} \\
     \ell_{21}\ell_{41} + s_2 \ell_{42} &      \ell_{21}\ell_{51} + s_2 \ell_{52} &      \ell_{21}\ell_{61} + s_2 \ell_{62} \\
\ell_{31}\ell_{41} + \ell_{32}\ell_{42} + s_3 \ell_{43} & \ell_{31}\ell_{51} + \ell_{32}\ell_{52} + s_3 \ell_{53} & \ell_{31}\ell_{61} + \ell_{32}\ell_{62} + s_3 \ell_{63}
\end{array}\right).
\end{equation}
Note that $\boldsymbol{\Sigma}_{pq}$ contains only spatial scales $s_1, s_2, s_3$ and bounded correlations $\ell_{ij}$ -- the query dimension scales $s_4, s_5, s_6$ do \textit{not} appear. This occurs because displacement describes how spatial position changes with query deviation, which should scale with the Gaussian's \textit{spatial} extent, not its query dimension spread.

In general, each element of $\boldsymbol{\Sigma}_{pq}$ can be expressed as:
\begin{equation}
[\boldsymbol{\Sigma}_{pq}]_{ij} = \sum_{k=1}^{i} L_{ik} L_{(3+j),k} = \sum_{k=1}^{i-1} \ell_{ik} \ell_{(3+j),k} + s_i \cdot \ell_{(3+j),i}.
\end{equation}

Cross-terms from spatial correlations $\ell_{ik}$ (where $k < i$) accumulate in the displacement. The displacement in row $i$ satisfies:
\begin{equation}
\|\boldsymbol{\Sigma}_{pq}^{(i)}\|_2 \leq \sqrt{C}(i-1 + s_i).
\end{equation}
For large $s_i$, this approximates $s_i \sqrt{C}$, but the constant offset $(i-1)\sqrt{C}$ can dominate for small Gaussians. This bound also introduces an asymmetry across spatial dimensions: even when $s_1 = s_2 = s_3$, row $i$ accumulates more cross-terms than row 1, allowing arbitrarily different displacement magnitudes depending on the spatial dimension ordering. While N-DGS uses zero initialization of $\ell_{ij}$ to ensure displacement initially scales with $s_i$, training can learn non-zero correlations that introduce both the offset and this asymmetry.

\textbf{Our method: Explicit coupling via scalar scaling.} We directly parameterize $\mathbf{V}_{pq} = \bar{s} \boldsymbol{\Theta}_{pq}$, where $\bar{s} = \text{mean}(s_x, s_y, s_z)$ is the average spatial scale and $\boldsymbol{\Theta}_{pq} \in \mathbb{R}^{3 \times C}$ are learnable parameters normalized to unit Frobenius norm ($\|\boldsymbol{\Theta}_{pq}\|_F = 1$). This gives:
\begin{equation}
\|\mathbf{V}_{pq}^{(i)}\|_2 = \bar{s} \|\boldsymbol{\Theta}_{i,:}\|_2 \leq \bar{s}.
\end{equation}
Each element in row $i$ scales uniformly with $\bar{s}$ with no cross-terms from other spatial dimensions, ensuring displacement scales with the average Gaussian size even for tiny Gaussians. The unit Frobenius norm constraint ensures bounded displacements while allowing flexible direction learning.

\textbf{Key difference.} Both methods bound displacement by spatial scale, but differ fundamentally:
\begin{itemize}[leftmargin=*]
\item \textbf{N-DGS}: Displacement emerges \textit{implicitly} from Cholesky structure with bound $\sqrt{C}(i-1 + s_i)$. Cross-terms can allow disproportionately large displacements for small Gaussians and introduce asymmetry across spatial dimensions. For highly anisotropic Gaussians (e.g., $s_x \gg s_y, s_z$), displacement along each axis $i$ is constrained by its individual scale $s_i$, allowing excessive motion along large axes while limiting movement along smaller dimensions.
\item \textbf{\modelname{}}: Displacement is \textit{explicitly} parameterized with tight bound $\bar{s}$, ensuring proper scaling across all Gaussian sizes with uniform treatment of all spatial dimensions. Critically, using the mean $\bar{s}$ enables balanced displacement control: it allows sufficient movement along smaller dimensions while preventing excessive motion along larger axes, ensuring the Gaussian can move uniformly in all directions regardless of anisotropy.
\end{itemize}

Empirically, removing spatial scaling (ablation: \textit{d}GS w/o $\bar{s}$) degrades PSNR by 2.59 dB on NeRF Synthetic (Table~\ref{tab:ablation}), confirming that displacement scaling is critical. Our direct parameterization provides tighter bounds, avoids dimension-ordering artifacts, and enables learnable per-dimension coupling strength control via $\Lambda$.


\section{Detailed Per-Scene Results}

This section provides comprehensive per-scene breakdowns for all experiments. Table~\ref{tab:detailed_combined} presents quality metrics (PSNR, SSIM, LPIPS) for each scene across all five datasets, while Table~\ref{tab:efficiency_combined} reports efficiency metrics (Gaussian count, training time, rendering FPS). These detailed results demonstrate the consistency of our method's performance across diverse scene types and complexity levels.

% Combined Detailed Quantitative Comparison Table
\begin{table*}[bp]
  \centering
  \caption{\small{Detailed quantitative comparison across all datasets at 30K iterations. N-DGS, dGS-O, and dGS are compared on static benchmarks (NeRF Synthetic, Mip-NeRF 360, 6DGS PBR) and dynamic benchmarks (7DGS PBR, D-NeRF).}} %Note, NeRF Synthetic and 7DGS PBR use decoupled opacity and position.
  \resizebox{0.90\linewidth}{!}{
        \begin{tabular}{@{\hskip 2pt}cc|@{\hskip 2pt}l@{\hskip 2pt}|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}}

    \toprule

    \multicolumn{2}{@{\hskip 2pt}c|@{\hskip 2pt}}{\multirow{2}[4]{*}{Dataset}} & \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{2}[4]{*}{Scene}}
        & \multicolumn{3}{c|@{\hskip 2pt}}{N-DGS}
        & \multicolumn{3}{c|@{\hskip 2pt}}{\textit{d}GS-O (\textbf{Ours})}
        & \multicolumn{3}{@{\hskip 2pt}c}{\textit{d}GS (\textbf{Ours})} \\
\cmidrule{4-12}
          &&       & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{PSNR$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{SSIM$\uparrow$}
                  & \multicolumn{1}{c|@{\hskip 2pt}}{LPIPS$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{PSNR$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{SSIM$\uparrow$}
                  & \multicolumn{1}{c|@{\hskip 2pt}}{LPIPS$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{PSNR$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{SSIM$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{LPIPS$\downarrow$} \\
    \midrule

\multicolumn{1}{@{\hskip 2pt}c}{\multirow{30}[4]{*}{\begin{sideways}STATIC SCENES\end{sideways}}}

    % ---------------- nerf_synthetic ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{9}[4]{*}{\begin{sideways}NeRF Synthetic\end{sideways}}}
    & \texttt{chair}     & 35.75 & 0.986 & 0.013 & \bf 36.36 & \bf 0.988 & \bf 0.012 &  36.26 & \bf 0.988 & \bf 0.012 \\
    & & \texttt{drums}     & 26.75 & 0.956 & 0.036 & 26.80 & 0.957 & 0.035 & \bf 26.92 & \bf 0.958 & \bf 0.034 \\
    & & \texttt{ficus}     & 33.54 & 0.984 & 0.014 & \bf 35.30 & \bf 0.988 & \bf 0.011 & 35.03 & 0.988 & \bf 0.011 \\
    & & \texttt{hotdog}    & 38.10 & 0.985 & 0.022 & \bf 38.13 & \bf 0.986 & \bf 0.020 & \bf 38.15 & \bf 0.986 & \bf 0.020 \\
    & & \texttt{lego}      & 35.63 & 0.981 & 0.019 & \bf 35.87 & \bf 0.982 & \bf 0.017 & 35.79 & \bf 0.982 & \bf 0.017 \\
    & & \texttt{materials} & 30.78 & 0.967 & 0.033 & 31.00 & 0.967 & 0.030 & \bf 31.76 & \bf 0.972 & \bf 0.027 \\
    & & \texttt{mic}       & 35.63 & 0.992 & 0.006 & 36.12 & \bf 0.993 & \bf 0.005 & \bf 36.25 & \bf 0.993 & \bf 0.005 \\
    & & \texttt{ship}      & 30.85 & 0.906 & 0.112 & \bf 31.16 & \bf 0.908 & 0.107 & 31.13 & \bf 0.908 & \bf 0.106 \\
\cmidrule{3-12}
    & & \texttt{avg}
      & 33.38
      & 0.970
      & 0.032
      & 33.84
      & 0.971
      & 0.030
      & \bf 33.91
      & \bf 0.972
      & \bf 0.029 \\
    \cmidrule{2-12}

    % ---------------- mipnerf360 ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{10}[4]{*}{\begin{sideways}Mip-NeRF 360\end{sideways}}}
        & \texttt{bicycle}   & 22.33 & 0.554 & 0.447 & 23.49 & 0.644 & 0.327 & \bf 23.51 & \bf 0.648 & \bf 0.318 \\
        & & \texttt{bonsai}    & 33.20 & 0.940 & 0.208 & 33.22 & 0.949 & 0.180 & \bf 33.46 & \bf 0.951 & \bf 0.175 \\
        & & \texttt{counter}   & 29.26 & 0.895 & 0.236 & 29.73 & 0.917 &  0.185 & \bf 29.93 & \bf 0.920 & \bf 0.181 \\
        & & \texttt{flowers}   & 18.90 & 0.409 & 0.507 &  19.61 & \bf 0.453 & \bf 0.443 & \bf 19.71 & 0.452 & 0.455 \\
        & & \texttt{garden}    & 26.19 & 0.809 & 0.205 & 26.77 & 0.842 &  0.145 & \bf 27.13 & \bf 0.849 & \bf 0.137 \\
        & & \texttt{kitchen}   & 31.17 & 0.922 & 0.137 & 31.72 & 0.930 &  0.121 & \bf 32.04 & \bf 0.933 & \bf 0.116 \\
        & & \texttt{room}      & 31.60 & 0.910 & 0.243 & 31.83 & 0.927 & 0.199 & \bf 32.25 & \bf 0.929 & \bf 0.195 \\
        & & \texttt{stump}     & 23.32 & 0.638 & 0.413 & \bf 26.01 &  0.750 & \bf 0.275 & 26.00 & \bf 0.742 & 0.295 \\
        & & \texttt{treehill}  & 21.66 & 0.571 & 0.448 & 22.21 & 0.603 & 0.393 & \bf 22.31 & \bf 0.606 & \bf 0.389 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & 26.40
          & 0.739
          & 0.316
          & 27.18
          & 0.779
          & 0.252
          & \bf 27.37
          & \bf 0.781
          & \bf 0.251 \\
\cmidrule{2-12}

  % ---------------- tandt_pbr ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{8}[4]{*}{\begin{sideways}6DGS PBR\end{sideways}}}
    & \texttt{bunny} & 40.87 & 0.993 & 0.045 & 43.44 & 0.993 & 0.042 & \bf 44.14 & \bf 0.993 & \bf 0.040 \\
    & & \texttt{cloud}     & 41.93 & 0.991 & 0.052 & 43.47 & 0.992 & 0.048 & \bf 43.53 & \bf 0.992 & \bf 0.045 \\
    & & \texttt{dragon}    & \bf 35.69 & \bf 0.947 & \bf 0.039 & 34.42 & 0.914 & 0.051 &  36.68 &  0.947 & 0.039 \\
    & & \texttt{explosion} & \bf 40.97 & \bf 0.989 & \bf 0.037 & 37.81 & 0.983 & 0.041 & 38.95 & 0.985 & 0.039 \\
    & & \texttt{ct-scan}     & 33.48 & 0.964 & 0.057 & 31.81 & 0.956 & 0.067 & \bf 34.62 & \bf 0.970 & \bf 0.048 \\
    & & \texttt{smoke}     & \bf 40.02 & \bf 0.990 & \bf 0.041 & 38.20 & 0.989 & 0.049 & 39.28 & \bf 0.990 & 0.047 \\
    & & \texttt{suzanne}   & 26.81 & 0.929 & \bf 0.106 & 26.65 & 0.927 & 0.112 & \bf 27.03 & \bf 0.930 & 0.107 \\
\cmidrule{3-12}
    & & \texttt{avg}
      & 37.11
      & 0.971
      & 0.054
      & 36.54
      & 0.965
      & 0.059
      & \bf 37.75
      & \bf 0.972
      & \bf 0.052 \\

    \midrule
    \midrule

\multicolumn{1}{@{\hskip 2pt}c}{\multirow{16}[4]{*}{\begin{sideways}DYNAMIC SCENES\end{sideways}}}

% & \multirow{7}{*}{\begin{sideways}7DGS PBR\end{sideways}}
%     & \texttt{cloud}    & \bf 30.31 & \bf 0.956 & \bf 0.072 & 29.34 & 0.955 & 0.075 & 29.60 & 0.955 & 0.074 \\
%     & & \texttt{dust}     & \bf 39.30 & \bf 0.959 & \bf 0.031 & 36.87 & 0.955 & 0.037 & 36.63 & 0.954 & 0.037 \\
%     & & \texttt{flame}    & \bf 33.90 & \bf 0.945 & \bf 0.054 & 30.38 & 0.933 & 0.066 & 30.32 & 0.932 & 0.066 \\
%     & & \texttt{heart}   & \bf 36.09 & \bf 0.987 & \bf 0.018 & 35.14 & 0.985 & 0.021 & 35.25 & 0.985 & 0.020 \\
%     & & \texttt{heart\_1600}   & \bf 32.15 & \bf 0.966 & \bf 0.048 & 31.29 & 0.962 & 0.052 & 31.36 & 0.962 & 0.051 \\
%     & & \texttt{suzanne}  & \bf 27.53 & \bf 0.947 & \bf 0.061 & 27.20 & 0.940 & 0.069 & 27.19 & 0.940 & 0.072 \\
%     \cmidrule{3-12}
%         & & \texttt{avg}
%           & \bf 33.21
%           & \bf 0.960
%           & \bf 0.047
%           & 31.70
%           & 0.955
%           & 0.053
%           & 31.72
%           & 0.955
%           & 0.053 \\
%     \cmidrule{2-12}


& \multirow{7}{*}{\begin{sideways}7DGS PBR\end{sideways}}
    & \texttt{cloud}    & 29.28 & 0.955 & 0.075 & 29.34 & 0.955 & 0.075 & \bf 29.82 & \bf 0.955 & \bf 0.074 \\
    & & \texttt{dust}     & 36.85 & 0.955 & 0.038 & 36.87 & 0.955 &  0.037 & \bf 38.39 & \bf 0.958 & \bf 0.031 \\
    & & \texttt{flame}    & \bf 31.64 & \bf 0.937 & \bf 0.062 & 30.38 & 0.933 & 0.066 & 31.52 & 0.936 & 0.060 \\
    & & \texttt{heart1}   & 34.66 & 0.983 & 0.023 & \bf 35.14 & \bf 0.985 & 0.021 & 35.25 & \bf 0.986 & \bf 0.018 \\
    & & \texttt{heart2}   & 30.99 & 0.959 & 0.057 & \bf 31.29 & \bf 0.962 & 0.052 & 31.01 & \bf 0.962 & \bf 0.049 \\
    & & \texttt{suzanne}  & 27.09 & \bf 0.941 & 0.072 & \bf 27.20 & 0.940 & \bf 0.069 & 27.07 & 0.940 & 0.070 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & 31.75
          & 0.955
          & 0.055
          & 31.70
          & 0.955
          & 0.053
          & \bf 32.18
          & \bf 0.956
          & \bf 0.050 \\
    \cmidrule{2-12}

& \multirow{9}{*}{\begin{sideways}D-NeRF\end{sideways}}
        & \texttt{b.balls}  & \bf 34.00 & \bf 0.981 & 0.025 & 32.52 & 0.978 & 0.028 &  33.59 & \bf 0.981 & \bf 0.023 \\
        & & \texttt{h.warrior}& 32.76 & 0.932 & 0.093 & 33.54 & 0.941 & 0.085 & \bf 34.64 & \bf 0.947 & \bf 0.073 \\
        & & \texttt{hook}     & 31.15 & 0.960 & 0.044 & 28.55 & 0.937 & 0.062 & \bf 32.94 & \bf 0.970 & \bf 0.030 \\
        & & \texttt{j.jacks}  & \bf 31.22 & \bf 0.967 & 0.043 & 28.62 & 0.952 & 0.054 & 30.86 & 0.965 & \bf 0.039 \\
        & & \texttt{lego}     & \bf 28.33 & \bf 0.943 & 0.055 & 27.13 & 0.939 &  0.055 & 28.06 & 0.942 & \bf 0.054 \\
        & & \texttt{mutant}   & 39.38 & 0.993 & 0.007 & 34.55 & 0.981 & 0.019 & \bf 39.84 & \bf 0.993 & \bf 0.006 \\
        & & \texttt{standup}  & 38.50 & 0.988 & 0.013 & 33.02 & 0.972 & 0.029 & \bf 38.76 & \bf 0.990 & \bf 0.009 \\
        & & \texttt{trex}     & 29.68 & 0.978 & 0.023 & 29.42 & 0.975 & 0.022 & \bf 30.08 & \bf 0.978 & \bf 0.021 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & 33.13
          & 0.968
          & 0.038
          & 30.92
          & 0.959
          & 0.044
          & \bf 33.59
          & \bf 0.971
          & \bf 0.032 \\
    \bottomrule
    \end{tabular}
    }
  \label{tab:detailed_combined}
\end{table*}



% Combined Detailed Efficiency Comparison Table
\begin{table*}[tbp]
  \centering
  \caption{\small{Detailed efficiency comparison across all datasets at 30K iterations. N-DGS, \textit{d}GS-O, and \textit{d}GS are compared showing number of Gaussians (K), training time (minutes), and rendering FPS on static benchmarks (NeRF Synthetic, Mip-NeRF 360, 6DGS PBR) and dynamic benchmarks (7DGS PBR, D-NeRF).}} %Note, NeRF Synthetic and 7DGS PBR use decoupled opacity and position.
  \resizebox{0.96\linewidth}{!}{
        \begin{tabular}{@{\hskip 2pt}cc|@{\hskip 2pt}l@{\hskip 2pt}|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r|@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}r@{\hskip 2pt}}

    \toprule

    \multicolumn{2}{@{\hskip 2pt}c|@{\hskip 2pt}}{\multirow{2}[4]{*}{Dataset}} & \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{2}[4]{*}{Scene}}
        & \multicolumn{3}{c|@{\hskip 2pt}}{N-DGS}
        & \multicolumn{3}{c|@{\hskip 2pt}}{\textit{d}GS-O (\textbf{Ours})}
        & \multicolumn{3}{@{\hskip 2pt}c}{\textit{d}GS (\textbf{Ours})} \\
\cmidrule{4-12}
          &&       & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{\#Gauss$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{Time$\downarrow$}
                  & \multicolumn{1}{c|@{\hskip 2pt}}{FPS$_{\text{CUDA}}$$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{\#Gauss$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{Time$\downarrow$}
                  & \multicolumn{1}{c|@{\hskip 2pt}}{FPS$_{\text{CUDA}}$$\uparrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{\#Gauss$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{Time$\downarrow$}
                  & \multicolumn{1}{@{\hskip 2pt}c@{\hskip 2pt}}{FPS$_{\text{CUDA}}$$\uparrow$} \\
    \midrule

\multicolumn{1}{@{\hskip 2pt}c}{\multirow{30}[4]{*}{\begin{sideways}STATIC SCENES\end{sideways}}}

    % ---------------- nerf_synthetic ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{9}[4]{*}{\begin{sideways}NeRF Synthetic\end{sideways}}}
    & \texttt{chair}     & \bf 178 & 6.11 & 635.27 & 206 & \bf 5.54 & \bf 713.54 & 199 & 5.86 & 677.38 \\
    & & \texttt{drums}     & \bf 193 & 6.19 & 476.51 & 239 & \bf 5.62 & \bf 683.86 & 242 & 6.04 & 646.90 \\
    & & \texttt{ficus}     & \bf 117 & 4.58 & 840.56 & 147 & \bf 4.27 & \bf 973.46 & 136 & 4.53 & 900.59 \\
    & & \texttt{hotdog}    & 102 & 6.04 & 871.45 & \bf 98 & \bf 5.56 & \bf 982.05 & 104 & 6.02 & 876.19 \\
    & & \texttt{lego}      & \bf 186 & 7.06 & 480.97 & 220 & \bf 5.67 & \bf 697.44 & 225 & 6.23 & 669.15 \\
    & & \texttt{materials} & \bf 136 & 5.38 & 767.74 & 182 & \bf 5.09 & 759.30 & 180 & 5.54 & 689.29 \\
    & & \texttt{mic}       & \bf 188 & 5.84 & 486.29 & 260 & 6.21 & \bf 625.05 & 260 & 11.17 & 594.08 \\
    & & \texttt{ship}      & 271 & 9.60 & 365.77 & \bf 256 & \bf 8.68 & \bf 605.15 & 259 & 9.28 & 585.17 \\
\cmidrule{3-12}
    & & \texttt{avg}
      & \bf 171
      & 6.35
      & 615.57
      & 201
      & \bf 5.83
      & \bf 754.98
      & 201
      & 6.84
      & 704.84 \\
    \cmidrule{2-12}

    % ---------------- mipnerf360 ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{10}[4]{*}{\begin{sideways}Mip-NeRF 360\end{sideways}}}
        & \texttt{bicycle}   & \bf 2243 & \bf 32.54 & 70.31 & 4452 & 33.46 & \bf 144.01 & 4644 & 38.30 & 118.20 \\
        & & \texttt{bonsai}    & \bf 751 & 23.49 & 186.80 & 1048 & \bf 20.02 & \bf 419.45 & 1091 & 21.77 & 359.76 \\
        & & \texttt{counter}   & \bf 555 & 22.79 & 237.17 & 880 & \bf 22.41 & \bf 430.22 & 902 & 23.98 & 380.55 \\
        & & \texttt{flowers}   & \bf 1547 & 26.28 & 100.84 & 2587 & \bf 24.43 & \bf 237.63 & 2679 & 26.27 & 196.76 \\
        & & \texttt{garden}    & \bf 2674 & 38.90 & 58.64 & 3552 & \bf 29.36 & \bf 174.37 & 3767 & 33.70 & 141.57 \\
        & & \texttt{kitchen}   & \bf 770 & 27.28 & 178.89 & 1148 & \bf 24.95 & \bf 354.03 & 1175 & 26.41 & 312.45 \\
        & & \texttt{room}      & \bf 792 & 24.62 & 177.95 & 1286 & \bf 23.81 & \bf 351.34 & 1356 & 25.75 & 298.32 \\
        & & \texttt{stump}     & \bf 644 & 21.29 & 212.22 & 3272 & \bf 27.56 & 199.15 & 3343 & 30.01 & 165.52 \\
        & & \texttt{treehill}  & \bf 1481 & 26.62 & 104.35 & 3042 & \bf 24.22 & \bf 203.51 & 3209 & 27.64 & 167.23 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & \bf 1273
          & 27.09
          & 147.46
          & 2363
          & \bf 25.58
          & \bf 279.30
          & 2463
          & 28.21
          & 237.82 \\
\cmidrule{2-12}

  % ---------------- tandt_pbr ----------------
& \multicolumn{1}{c|@{\hskip 2pt}}{\multirow{8}[4]{*}{\begin{sideways}6DGS PBR\end{sideways}}}
    & \texttt{bunny} & \bf 24 & 33.69 & \bf 718.47 & 63 & 42.62 & 689.88 & 62 & 45.03 & 642.86 \\
    & & \texttt{cloud}     & \bf 17 & 32.61 & 597.10 & 62 & 39.08 & \bf 700.35 & 60 & 41.31 & 654.93 \\
    & & \texttt{dragon}    & \bf 93 & 12.76 & \bf 855.89 & 118 & 12.87 & 787.64 & 134 & 13.28 & 753.36 \\
    & & \texttt{explosion} & \bf 35 & 22.13 & \bf 875.48 & 72 & 29.54 & 813.28 & 73 & 40.77 & 753.18 \\
    & & \texttt{ct-scan}     & \bf 171 & 6.57 & 689.47 & 221 & \bf 6.44 & \bf 796.54 & 299 & 7.45 & 695.69 \\
    & & \texttt{smoke}     & \bf 52 & 46.48 & \bf 608.02 & 72 & 55.67 & 568.30 & 74 & 58.15 & 547.93 \\
    & & \texttt{suzanne}   & \bf 201 & 33.16 & 404.34 & 221 & \bf 28.80 & \bf 518.17 & 233 & 30.58 & 526.10 \\
\cmidrule{3-12}
    & & \texttt{avg}
      & \bf 85
      & 26.77
      & 678.40
      & 118
      & 30.72
      & \bf 696.31
      & 134
      & 33.80
      & 653.44 \\

    \midrule
    \midrule

\multicolumn{1}{@{\hskip 2pt}c}{\multirow{16}[4]{*}{\begin{sideways}DYNAMIC SCENES\end{sideways}}}

& \multirow{7}{*}{\begin{sideways}7DGS PBR\end{sideways}}
    & \texttt{cloud}    & \bf 48 & 76.87 & 950.04 & 63 & \bf 75.47 & \bf 953.38 & 70 & 76.11 & 892.36 \\
    & & \texttt{dust}     & \bf 16 & 13.11 & 1052.85 & 23 & \bf 12.06 & \bf 1165.85 & 30 & 16.89 & 990.11 \\
    & & \texttt{flame}    & \bf 22 & \bf 14.91 & \bf 1063.87 & 23 & 16.87 & 1015.92 & 40 & 19.30 & 926.17 \\
    & & \texttt{heart1}   & \bf 140 & 30.75 & 593.96 & 182 & \bf 28.75 & \bf 697.33 & 188 & 30.08 & 639.83 \\
    & & \texttt{heart2}   & \bf 181 & 22.75 & 432.58 & 237 & \bf 20.12 & \bf 729.08 & 261 & 22.60 & 662.23 \\
    & & \texttt{suzanne}  & \bf 564 & 70.47 & 165.70 & 837 & \bf 54.23 & \bf 449.89 & 767 & 56.75 & 411.89 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & \bf 162
          & 38.14
          & 709.83
          & 227
          & \bf 34.58
          & \bf 835.24
          & 226
          & 36.96
          & 753.77 \\
    \cmidrule{2-12}

& \multirow{9}{*}{\begin{sideways}D-NeRF\end{sideways}}
        & \texttt{b.balls}  & 179 & 22.73 & 398.65 & \bf 103 & \bf 16.09 & \bf 770.99 & 128 & 17.72 & 636.37 \\
        & & \texttt{h.warrior}& \bf 13 & 8.25 & \bf 1151.03 & 17 & \bf 7.87 & 1141.34 & 18 & 9.23 & 1068.65 \\
        & & \texttt{hook}     & \bf 37 & \bf 9.71 & \bf 1042.00 & 55 & 11.42 & 1027.26 & 75 & 14.58 & 865.61 \\
        & & \texttt{j.jacks}  & \bf 25 & \bf 9.61 & 1069.25 & 31 & 10.67 & \bf 1101.29 & 71 & 14.48 & 899.56 \\
        & & \texttt{lego}     & \bf 114 & 21.09 & 583.20 & 143 & \bf 17.93 & \bf 700.47 & 125 & 19.28 & 685.07 \\
        & & \texttt{mutant}   & \bf 63 & 14.74 & 829.30 & 67 & \bf 13.27 & \bf 930.20 & 79 & 15.47 & 810.36 \\
        & & \texttt{standup}  & \bf 21 & 9.71 & 1074.73 & 26 & \bf 9.28 & \bf 1140.56 & 51 & 14.00 & 867.39 \\
        & & \texttt{trex}     & \bf 101 & \bf 16.77 & 667.96 & 118 & 19.09 & \bf 778.03 & 134 & 19.69 & 714.27 \\
    \cmidrule{3-12}
        & & \texttt{avg}
          & \bf 69
          & 14.08
          & 852.01
          & 70
          & \bf 13.20
          & \bf 948.77
          & 85
          & 15.56
          & 818.41 \\
    \bottomrule
    \end{tabular}
    }
  \label{tab:efficiency_combined}
\end{table*}


\end{document}
